---
title: "analysis"
author: "Sara Irish"
date: "`r Sys.Date()`"
output: html_document
---
#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
  library(glmmTMB)
  library(car)
  library(popbio)
  library(ggplot2)
  library(ggbeeswarm)
  library(shades)
  library(RColorBrewer)
  library(DHARMa)
  library(Rmisc)
  library(dabestr)
  library(magrittr)
  library(tidyr)
  library(dplyr)
  library(ggeffects)
  library(lme4)
  library(optimx)
  library(nloptr)
  library(dfoptim)
  library(data.table)
  library(pscl)
  library(ggThemeAssist)
  library(DHARMa)
  library(emmeans)
  library(ggpubr)
  library(patchwork)
  library(lubridate)
  library(survival)
  library(coxme) #for cox model
  library(ggplot2)
  library(purrr) #for forestplot
  library(ggpubr) #for forestplot
  library(survminer) #for forestplot
  library(AICcmodavg)
  library(permutes)
  library(performance)
  library(fields)
  library(grid)
  library(ggplotify)
  library(cowplot)
  library(eha)
  library(multcomp)
  library(usethis)
  library(forcats)
})

usethis::use_github()
#getwd()
#setwd("C:/Users/irish/OneDrive/Documents/R/SKN-1_IIS/SKN-1_IIS")
#24-hour 15 to 25C thermocycling experiments, HF = 25C first, CF = 15C first, LS = lifespan, rep = reproduction
thermo_HF_rep <- read.csv('data/Thermo_H_rep.csv')
thermo_HF_LS <- read.csv('data/Thermo_H_LS.csv')
thermo_CF_rep <- read.csv('data/Thermo_C_rep.csv')
thermo_CF_LS <- read.csv('data/Thermo_C_LS.csv') 

#SKN-1 mutants with and without daf-2 RNAi KD in 20C - LS = lifespan, rep = reproduction
SKN_20C_rep <- read.csv('data/SKN1_20C_repro.csv')
SKN_20C_LS <- read.csv('data/SKN1_20C_LS.csv')

#Lifespan experiment SKN-1 mutants with and without daf-2 RNAi KD in 15 and 25C, with additional treatment of daf-2 RNAi KD SKN-1 mutants starting with 24 hrs of opposite temperature to test effects of temperature on early development of SKN-1 mutants
SKN_15_25_LS <- read.csv('data/SKN_N2_LS_15_25.csv')

```



#SKN-1 in 20C
Reproduction and lifespan assays in SKN-1 mutants with daf-2 or ev (control) RNAi treatments in 20C

##Reproduction
```{r}

#put data into long form
skn_long <- SKN_20C_rep %>% 
  pivot_longer(
    cols = c(`D1`, `D2`, `D3`, `D4`, `D5`), 
    names_to = "Day",
    values_to = "value")

summarySE(thermo_HF_LS, groupvars = c('Strain','Treatment.ID'), measurevar = 'Age')


#select only columns needed
skn_long <- skn_long %>%
  dplyr::select(c('ID','Treatment','Day','value'))

#rename treatments
skn_long$Treatment <- as.factor(skn_long$Treatment)
levels(skn_long$Treatment) <- list('ev' = 'ev', 'daf-2' = 'daf')

#Get Day into a numeric form
#Change day to a numeric variable for analysis
skn_long$Day <- as.factor(skn_long$Day)
levels(skn_long$Day) <- list('1' = 'D1', '2' = 'D2', '3' = 'D3', '4' = 'D4', '5' = 'D5')
skn_long$Day <- as.numeric(skn_long$Day)

#set ev as reference level
skn_long$Treatment <- as.factor(skn_long$Treatment)
skn_long$Treatment <- relevel(skn_long$Treatment, ref = 'ev')

#Age-specific reproduction aanalysis with negative binomial GLMM
m1_SKN <- glmmTMB(value ~ Treatment * Day + (1|ID), family = nbinom2(), data = skn_long)
summary(m1_SKN)

#model diagnostics
sim_SKN1 <- simulateResiduals(m1_SKN, plot = T)
testDispersion(sim_SKN1)#underdispersed
testZeroInflation(sim_SKN1)#zero-inflated

#Try adding zero inflation formula
m2_SKN <- glmmTMB(value ~ Treatment*Day + (1|ID), ziformula = ~Day, family = nbinom2(), data = skn_long)
summary(m2_SKN)#won't converge, try removing interaction (appears nonsig in previous model)

#Remove interaction
m3_SKN <- glmmTMB(value ~ Treatment + Day + (1|ID), family = nbinom2(), ziformula = ~ Day, data = skn_long)
summary(m3_SKN)

#diagnostic
sim_SKN3 <- simulateResiduals(m3_SKN, plot = T)
testZeroInflation(sim_SKN3)
testDispersion(sim_SKN3)

### Try a poisson model
m4_SKN <- glmmTMB(value ~ Treatment * poly(Day,2) + (1|ID), family = poisson(), data = skn_long)
AIC(m1_SKN, m4_SKN)#poisson has MUCH higher AIC
summary(m4_SKN)
sim_m4 <- simulateResiduals(m4_SKN, plot = T)
testDispersion(sim_m4)#dispersion much better
testZeroInflation(sim_m4)#major ZI
#compare other models
AIC(m1_SKN, m2_SKN, m3_SKN, m4_SKN) #lowest AIC - but no dispersion issues

# **** Final model
### Try a poisson model with ziformula
m5_SKN <- glmmTMB(value ~ Treatment * poly(Day,2) + (1|ID), ziformula = ~Day, family = poisson(), data = skn_long)
summary(m5_SKN)
sim_m5 <- simulateResiduals(m5_SKN, plot = T)
testZeroInflation(sim_m5)
testDispersion(sim_m5)
Anova(m5_SKN, type = 'III')
AIC(m4_SKN, m5_SKN)

#to get anova results for full model
m5_SKN_anova <- glmmTMB(value ~ Treatment * I(Day^2)+Treatment*Day + (1|ID), ziformula = ~Day, family = poisson(), data = skn_long)
Anova(m5_SKN, type = 'III')


##############
#### LRS
##############

#sum up total number of offspring produced over experiment per individual
totalrep_SKN<-na.omit(as.data.frame.table(tapply(skn_long$value,list(skn_long$Treatment, skn_long$ID),sum)))

#rename columns
names(totalrep_SKN)<-c("Treatment", "Replicate", "Totrep")#

#Look at distribution
  hist(totalrep_SKN$Totrep)#pretty skewed - but try LM

#Try running basic linear model
SKN_LRS_m1 <- lm(value ~ Treatment, data = skn_long)
summary(SKN_LRS_m1)
SKN_LRS_sim1 <- simulateResiduals(SKN_LRS_m1, plot = T)

#Try a Gamma GLM
SKN_LRS_m2 <- glm(Totrep ~ Treatment, data = totalrep_SKN, family = Gamma(link = 'log'))
summary(SKN_LRS_m2)
SKN_LRS_sim2 <- simulateResiduals(SKN_LRS_m2, plot = T)
Anova(SKN_LRS_m2, type = 'II')
#Try link = inverse
SKN_LRS_m3 <- glm(Totrep ~ Treatment, data = totalrep_SKN, family = Gamma(link = 'inverse'))
summary(SKN_LRS_m3)
SKN_LRS_sim3 <- simulateResiduals(SKN_LRS_m3, plot = T)
Anova(SKN_LRS_m3, type = 'II')

totalrep_SKN$Treatment <- as.factor(totalrep_SKN$Treatment)

###########
#lambda#
##########

#### Lambda
#start fresh
skn_long <- SKN_20C_rep %>% 
  pivot_longer(
    cols = c(`D1`, `D2`, `D3`, `D4`, `D5`), 
    names_to = "Day",
    values_to = "value")

str(skn_long)
#keep only columns needed
skn_long <- skn_long %>%
  dplyr::select(c('ID', 'Strain','Treatment','Day','value'))
str(skn_long)


#spread data back out
skn_wide <- spread(skn_long, Day, value)


skn_wide <- na.omit(skn_wide)

#calculate Lambda for each individual
L <- matrix(nrow = nrow(skn_wide), ncol = 4)

for (i in 1:nrow(skn_wide)){
  Les <- matrix(0, ncol = 7, nrow = 7)
  diag(Les[-1,]) <- rep(1, 6) # add the 1s for survival probability diagonally
  Fert <- c(0,0, as.numeric(as.vector(skn_wide[i,][4:8]))) #the columns in data that has the reproductive counts
  Fert[is.na(Fert)] <- 0 #makes all NAs into 0s
  Les[1,] <- c(Fert)
  class(Les) <- "leslie.matrix"
  Lambda <- popbio::eigen.analysis(Les)$lambda1
  L[i, 1:4] <- c(paste0(skn_wide$ID[i]), paste0(skn_wide$Strain[i]), paste0(skn_wide$Treatment[i]), Lambda)
  
}

#rename columns
colnames(L)<-c("ID", "Strain", "Treatment", "Lambda")

#make L a data frame
skn_lam<-as.data.frame(L)

#make sure it's numeric
skn_lam$Lambda<-as.numeric(as.character(skn_lam$Lambda))
skn_lam$Treatment <- as.factor(skn_lam$Treatment)

skn_lam$Treatment <- revalue(skn_lam$Treatment, c('ev' = 'control', 'daf' = 'daf-2'))
skn_lam$Treatment <- relevel(skn_lam$Treatment, ref = 'control')

#analysis 
skn_lam_mod <- lm(Lambda ~ Treatment, data = skn_lam)
summary(skn_lam_mod)
Anova(skn_lam_mod, type = 'II')


#get means
skn_mean_lam <- summarySE(data = skn_lam, measurevar = 'Lambda', groupvars = c('Treatment'))
skn_mean_lam



```

##Lifespan
```{r}

#Get rid of NAs
SKN_20C_LS <- na.omit(SKN_20C_LS)

#Rename daf to daf-2
SKN_20C_LS$Treatment.ID <- revalue(SKN_20C_LS$Treatment.ID, c('ev' = 'ev', 'daf' = 'daf-2'))

#Make sure Treatment is a factor
SKN_20C_LS$Treatment.ID <- as.factor(SKN_20C_LS$Treatment.ID)

#Make ev the baseline level
SKN_20C_LS$Treatment.ID <- relevel(SKN_20C_LS$Treatment.ID, ref = 'ev')

#Check factor levels of treatment
levels(SKN_20C_LS$Treatment.ID)

#Get rid of empty levels
SKN_20C_LS$Treatment.ID <- droplevels(SKN_20C_LS$Treatment.ID)

#Plate.ID as a factor not a numerical
SKN_20C_LS$Plate.ID <- as.factor(SKN_20C_LS$Plate.ID)

#Mixed effects cox model for a effect of daf-2 on SKN-1 mutant lifespan
cox_SKN_20C <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = SKN_20C_LS)
summary(cox_SKN_20C)
Anova(cox_SKN_20C, type = 'II')

#test model
cox.zph(cox_SKN_20C)#OK

#Exclude matricides
SKN_20C_LS_NM <- SKN_20C_LS %>%
  filter(Cause != 'M')

#Cox mixed effects model
cox_SKN_20C_NM <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = SKN_20C_LS_NM)
summary(cox_SKN_20C_NM)
Anova(cox_SKN_20C_NM, type = 'II')
#Test model
cox.zph(cox_SKN_20C_NM)#OK

```


#15C first Thermocycling

Reproduction and lifespan assays following 24 hour thermo-cycling between 15 and 25C with 15C as starting temperature across four treatments: SKN-1 mutant + daf-2 RNAi, SKN-1 mutant + ev (control), N2 + daf-2 RNAi, N2 + ev

## Reproduction 
```{r}

#Dataframe into long form
thermo_c_long <- thermo_CF_rep %>% 
  pivot_longer(
    cols = c(`D1`,`D2`, `D3`, `D4`, `D5`,`D6`, `D7`,`D8`,`D9`), 
    names_to = "Day",
    values_to = "value")


#Remove unneeded columns
thermo_c_long <- thermo_c_long %>%
  dplyr::select(c('ID','Strain', 'Treatment','Day','value'))

#Remove NAs
thermo_c_long <- na.omit(thermo_c_long)

#Combine strain and treatment into one variable for plotting
thermo_c_long <- thermo_c_long %>%
  unite(Str_tr, c('Strain', 'Treatment'), remove = F)

#Set levels of StrainxTreatment
thermo_c_long$Str_tr <- factor(thermo_c_long$Str_tr, levels = c("N2_ev", "N2_daf", "SKN-1_ev", "SKN-1_daf"))

#change names of Str_tr
levels(thermo_c_long$Str_tr) <- list('N2 ev' = 'N2_ev', 'N2 daf-2' = 'N2_daf', 'SKN-1 ev' = 'SKN-1_ev', 'SKN-1 daf-2' = 'SKN-1_daf')

#Make sure Day is a factor
thermo_c_long$Day <- as.factor(thermo_c_long$Day)

#Change day to a numeric variable for analysis
levels(thermo_c_long$Day) <- list('1' = 'D1', '2' = 'D2', '3' = 'D3', '4' = 'D4', '5' = 'D5', '6' = 'D6', '7' = 'D7', '8' = 'D8', '9' = 'D9')
thermo_c_long$Day <- as.numeric(thermo_c_long$Day)


#Set ev as reference treatment level
thermo_c_long$Treatment <- as.factor(thermo_c_long$Treatment)
thermo_c_long$Treatment <- relevel(thermo_c_long$Treatment, ref = 'ev')

#Analyse age-specific repro - use glmmTMB negative binomial due to high level of zeroes
CF_m1 <- glmmTMB(value ~ Strain *Treatment*Day + (1|ID), family = nbinom1, data = thermo_c_long)

#Simulate residuals with DHARMa package to test for dispersion and zero-inflation issues
CF_sim1 <- simulateResiduals(CF_m1, plot = T)
testDispersion(CF_sim1)#underdispersed
testZeroInflation(CF_sim1)#zero-inflation

#Add zero-inflation parameter
CF_m2 <- glmmTMB(value ~ Strain * Treatment * Day + (1|ID), ziformula = ~ Day, family = 'nbinom1', data = thermo_c_long)

#SImulate residuals again to test model
CF_sim2 <- simulateResiduals(CF_m2, plot = T)
testDispersion(CF_sim2)
testZeroInflation(CF_sim2)
summary(CF_m2)

#Get chi-squared using type III Anova to account for interaction
Anova(CF_m2, type = 'III')


# Lifetime reproduction success

#Create dataframe containing Sum of total reproduction for each individual
totalrep_CF<-na.omit(as.data.frame.table(tapply(thermo_c_long$value,list(thermo_c_long$Strain, thermo_c_long$Treatment, thermo_c_long$Str_tr, thermo_c_long$ID),sum)))

#rename columns
names(totalrep_CF)<-c('Strain', 'Treatment', "Str_tr", "Replicate", "Totrep")

#Set factor levels for 
totalrep_CF$Str_tr <- factor(totalrep_CF$Str_tr, levels = c("N2 ev", "N2 daf-2", "SKN-1 ev", "SKN-1 daf-2"))

#Check distribution
hist(totalrep_CF$Totrep)#close to normal

#Analyse LRS with linear model
LRS_CF <- lm(Totrep ~ Strain * Treatment, data = totalrep_CF)

#Simulate residuals
LRS_CF_sim1 <- simulateResiduals(LRS_CF, plot = T)#looks good

summary(LRS_CF)#no effect, try removing interaction

LRS_CF2 <- lm(Totrep ~ Strain + Treatment, data = totalrep_CF)

#Simulate residuals
LRS_CF_sim2 <- simulateResiduals(LRS_CF2, plot = T)#OK

summary(LRS_CF2)#trend toward higher LRS in SKN-1

#Use type II anova to get chi-squared
Anova(LRS_CF2, type = 'II')

#Dataframe into long form to start fresh
thermo_c_long <- thermo_CF_rep %>% 
  pivot_longer(
    cols = c(`D1`,`D2`, `D3`, `D4`, `D5`,`D6`, `D7`,`D8`,`D9`), 
    names_to = "Day",
    values_to = "value")

#create a level order for 10 days
level_order <- c('D1', 'D2', 'D3', 'D4', 'D5','D6','D7','D8','D9')

#reorder levels so 10 is last
thermo_c_long$Day <- factor(thermo_c_long$Day, levels = level_order)


#spread data back out
thermo_c_wide <- spread(thermo_c_long, Day, value)
str(thermo_c_wide)

thermo_c_wide <- thermo_c_wide %>%
  dplyr::select(-c(Matricide))

thermo_c_wide <- na.omit(thermo_c_wide)

#calculate Lambda for each individual
L <- matrix(nrow = nrow(thermo_c_wide), ncol = 4)

for (i in 1:nrow(thermo_c_wide)){
  Les <- matrix(0, ncol = 11, nrow = 11)
  diag(Les[-1,]) <- rep(1, 10) # add the 1s for survival probability diagonally
  Fert <- c(0,0, as.numeric(as.vector(thermo_c_wide[i,][4:12]))) #the columns in data that has the reproductive counts
  Fert[is.na(Fert)] <- 0 #makes all NAs into 0s
  Les[1,] <- c(Fert)
  class(Les) <- "leslie.matrix"
  Lambda <- popbio::eigen.analysis(Les)$lambda1
  L[i, 1:4] <- c(paste0(thermo_c_wide$ID[i]), paste0(thermo_c_wide$Strain[i]), paste0(thermo_c_wide$Treatment[i]), Lambda)
  
}

#rename columns
colnames(L)<-c("ID", "Strain", "Treatment", "Lambda")

#make L a data frame
thermo_c_lam<-as.data.frame(L)

#make sure it's numeric
thermo_c_lam$Lambda<-as.numeric(as.character(thermo_c_lam$Lambda))
thermo_c_lam$Treatment <- as.factor(thermo_c_lam$Treatment)


#lambda analysis
hist(thermo_c_lam$Lambda)#there's an outlier
thermo_c_lam_m1 <- lm(Lambda ~ Strain * Treatment, data = thermo_c_lam)
summary(thermo_c_lam_m1)
sim_lam_m1 <- simulateResiduals(thermo_c_lam_m1, plot = T)#ok, outlier creates only minor issues

#try removing interaction
thermo_c_lam_m2 <- lm(Lambda ~ Strain + Treatment, data = thermo_c_lam)
summary(thermo_c_lam_m2)
sim_lam_m1 <- simulateResiduals(thermo_c_lam_m1, plot = T)
Anova(thermo_c_lam_m2, type = 'II')

```

## Lifespan
```{r}

#Create a combined variable for strain and treatment since forest plot can't do interactions
thermo_CF_LS <- thermo_CF_LS %>%
  unite(Str_tr, c('Strain', 'Treatment.ID'), remove = F)



#Set levels
thermo_CF_LS$Tr_str <- factor(thermo_CF_LS$Str_tr, levels = c("N2_ev", "N2_daf", "SKN-1_ev", "SKN-1_daf"))

#Set as factor
thermo_CF_LS$Str_tr <- as.factor(thermo_CF_LS$Str_tr)

#Set N2_ev as baseline level
thermo_CF_LS$Str_tr <- relevel(thermo_CF_LS$Str_tr, ref = 'N2_ev')

#Exclude unnatural causes of mortality, L = loss, W = wall, E = explosion
thermo_CF_LS <- thermo_CF_LS %>%
  filter(thermo_CF_LS$Cause != 'L' & thermo_CF_LS$Cause !='W' & thermo_CF_LS$Cause != 'E')

#Create dataframe excluding matricides
thermo_CF_LSNM <- thermo_CF_LS %>%
  filter(thermo_CF_LS$Cause != 'M')

#Sometimes NAs are produced creating an additional non-existent factor level - omit NAs first
thermo_CF_LS <- na.omit(thermo_CF_LS)
thermo_CF_LSNM <- na.omit(thermo_CF_LSNM)

#Drop non-existent factor level - or cox models get confused
thermo_CF_LS$Str_tr <- droplevels(thermo_CF_LS$Str_tr)
levels(thermo_CF_LS$Str_tr)
thermo_CF_LSNM$Str_tr <- droplevels(thermo_CF_LSNM$Str_tr)

#Mixed effects cox proportional hazards model to test effect of treatment on survival probability
CF_cox <- coxme(Surv(Age, Event) ~ Strain * Treatment.ID + (1|Plate.ID), data = thermo_CF_LS)
summary(CF_cox)
Anova(CF_cox, type = 'III')

#Test model 
cox.zph(CF_cox)#Strain * Treatment is fine
plot(cox.zph(CF_cox))

#Mixed effects cox proportional hazards model to test effect of treatment on survival prob
CF_NM_cox <- coxme(Surv(Age, Event) ~ Strain * Treatment.ID + (1|Plate.ID), data = thermo_CF_LSNM)
summary(CF_NM_cox)
Anova(CF_NM_cox, type = 'III')

#test model
plot(cox.zph(CF_NM_cox))
cox.zph(CF_NM_cox)#close p-values - combined with plot acceptable 


```


# 25C first thermocycling
Reproduction and lifespan assays following 24 hour thermo-cycling between 15 and 25C with 25C as starting temperature across four treatments: SKN-1 mutant + daf-2 RNAi, SKN-1 mutant + ev (control), N2 + daf-2 RNAi, N2 + ev

## Reproduction
```{r}

#Put dataframe into long form
thermo_h_long <- thermo_HF_rep %>% 
  pivot_longer(
    cols = c(`D1`,`D2`, `D3`, `D4`, `D5`,`D6`, `D7`,`D8`,`D9`), 
    names_to = "Day",
    values_to = "value")

#Keep only columns that are necessary
thermo_h_long <- thermo_h_long %>%
  dplyr::select(c('ID','Strain', 'Treatment','Day','value'))

#Get rid of NAs
thermo_h_long <- na.omit(thermo_h_long)

#Create a combined strain and treatment column for plotting
thermo_h_long <- thermo_h_long %>%
  unite(Str_tr, c('Strain', 'Treatment'), remove = F)

#Set factor levels for Strain x Treatment
thermo_h_long$Str_tr <- factor(thermo_h_long$Str_tr, levels = c("N2_ev", "N2_daf", "SKN-1_ev", "SKN-1_daf"))

#change names of Str_tr
levels(thermo_h_long$Str_tr) <- list('N2 ev' = 'N2_ev', 'N2 daf-2' = 'N2_daf', 'SKN-1 ev' = 'SKN-1_ev', 'SKN-1 daf-2' = 'SKN-1_daf')
#Make sure Day is a factor
thermo_h_long$Day <- as.factor(thermo_h_long$Day)

#Change day to a numeric variable for analysis
levels(thermo_h_long$Day) <- list('1' = 'D1', '2' = 'D2', '3' = 'D3', '4' = 'D4', '5' = 'D5', '6' = 'D6', '7' = 'D7', '8' = 'D8', '9' = 'D9')
thermo_h_long$Day <- as.numeric(thermo_h_long$Day)

#Analyse age-specific reproduction - glmmTMB with negative binomial family (will try nbinom2 to account for curvature at early days)

HF_m1 <- glmmTMB(value ~ Strain * Treatment * Day + (1|ID), family = nbinom2(), data = thermo_h_long)

#simulate residuals with DHARMa, test dispersion and zero inflation
HF_sim1 <- simulateResiduals(HF_m1, plot = T)
testDispersion(HF_sim1)# a bit underdispersed
testZeroInflation(HF_sim1)#very zero-inflated

#Try adding ZI paramter ~ Day (since 0's more likely to occur on certain days)
HF_m2 <- glmmTMB(value ~ Strain * Treatment * Day + (1|ID), family = nbinom2(), ziformula = ~Day, data = thermo_h_long)

#simualte residuals
HF_sim2 <- simulateResiduals(HF_m2, plot = T)
testDispersion(HF_sim2)#got a little worse
testZeroInflation(HF_sim2)#still very zero-inflated

#Try adding additional Treatment*Strain information into ZI parameter
HF_m3 <- glmmTMB(value ~ Strain * Treatment * Day + (1|ID), family = nbinom2(), ziformula = ~ Day * Strain * Treatment, data = thermo_h_long)

#simulate residuals
HF_sim3 <- simulateResiduals(HF_m3, plot= T)
testDispersion(HF_sim3)
testZeroInflation(HF_sim3)
summary(HF_m3)#not converging? all NA

#Try removing interactions from ziformula
HF_m4 <- glmmTMB(value ~ Strain * Treatment * Day + (1|ID), ziformula = ~ Day + Strain + Treatment, data = thermo_h_long, family = nbinom2())#won't converge

#Add OLRE - keep Day as formula
thermo_h_long <- tibble::rowid_to_column(thermo_h_long, "OLRE")
HF_m5 <- glmmTMB(value ~ Strain * Treatment * Day + (1|ID) + (1|OLRE), ziformula = ~ Day, family = nbinom2(), data = thermo_h_long)
summary(HF_m5)
HF_sim5 <- simulateResiduals(HF_m5, plot = T)
testDispersion(HF_sim5)
testZeroInflation(HF_sim5)

#Try including just strain in ziformula (because SKN-1s overall reproduced for longer)
HF_m6 <- glmmTMB(value ~ Strain * Treatment * Day + (1|ID) + (1|OLRE), ziformula = ~ Day + Strain, family = nbinom2(), data = thermo_h_long)
summary(HF_m6)#won't converge - previous model showed no signfiicant 3-way interaction - try removing

#Model without 3 way interaction
HF_m7 <- glmmTMB(value ~ Strain*Treatment + Strain * Day + Treatment *Day + (1|ID) + (1|OLRE), ziformula = ~Day + Strain, family = nbinom2(), data = thermo_h_long)
summary(HF_m7)#this converged
HF_sim7 <- simulateResiduals(HF_m7, plot = T)
testDispersion(HF_sim7)#better
testZeroInflation(HF_sim7)#better but still zero inflated

#Try adding 1 + Day as ziformula
HF_m8 <- glmmTMB(value ~ Strain*Treatment + Strain * Day + Treatment *Day + (1|ID) + (1|OLRE), ziformula = ~1+Day + Strain, family = nbinom2(), data = thermo_h_long)
summary(HF_m8)

HF_sim8 <- simulateResiduals(HF_m8, plot = T)
testDispersion(HF_sim8)#better
testZeroInflation(HF_sim8)#slightly better

#Try adding strain as interaction with day to ziformula
HF_m9 <- glmmTMB(value ~ Strain*Treatment + Strain * Day + Treatment *Day + (1|ID) + (1|OLRE), ziformula = ~1+Day * Strain, family = nbinom2(), data = thermo_h_long)
summary(HF_m9)#won't converge

#try adding strain additively to day in ziformula
HF_m10 <- glmmTMB(value ~ Strain*Treatment + Strain * Day + Treatment *Day + (1|ID) + (1|OLRE), ziformula = ~1+Day + Strain, family = nbinom1(), data = thermo_h_long)
summary(HF_m10)#doesn't converge


#Try using splines in ziformula
library(splines)
HF_m11 <- glmmTMB(value ~ Strain*Treatment + Strain*Day + Treatment*Day + (1|ID), ziformula =~ ns(Day, df = 3), family = nbinom2(), data = thermo_h_long)
summary(HF_m11)
HF_sim11 <- simulateResiduals(HF_m11, plot = T)
testDispersion(HF_sim11)
testZeroInflation(HF_sim11)
Anova(HF_m11, type = 'III')

#try df = 2 splines
HF_m11_2 <- glmmTMB(value ~ Strain*Treatment + Strain*Day + Treatment*Day + (1|ID), ziformula =~ ns(Day, df = 2), family = nbinom2(), data = thermo_h_long)

#try df = 4 splines
HF_m11_4 <- glmmTMB(value ~ Strain*Treatment + Strain*Day + Treatment*Day + (1|ID), ziformula =~ ns(Day, df = 4), family = nbinom2(), data = thermo_h_long)

#df = 3 is best - but makes little difference whether it's 2, 3, 4
AIC(HF_m11, HF_m11_2, HF_m11_4)


#Calculate total reproduction over experiment per individual
totalrep_HF<-na.omit(as.data.frame.table(tapply(thermo_h_long$value,list(thermo_h_long$Strain, thermo_h_long$Treatment, thermo_h_long$Str_tr, thermo_h_long$ID),sum)))

#rename columns
names(totalrep_HF)<-c('Strain','Treatment',"Str_tr", "Replicate", "Totrep")

#change factor levels
totalrep_HF$Str_tr <- factor(totalrep_HF$Str_tr, levels = c("N2 ev", "N2 daf-2", "SKN-1 ev", "SKN-1 daf-2"))



totalrep_HF$Treatment <- relevel(totalrep_HF$Treatment, ref = 'ev')

#Check distribution of LRS 
hist(totalrep_HF$Totrep)#normal

#Linear model to analyse total repro
LRS_HF_m1 <- lm(Totrep ~ Treatment*Strain, data = totalrep_HF)
summary(LRS_HF_m1)

#simulate residual with DHARMa package
LRS_HF_sim1 <- simulateResiduals(LRS_HF_m1, plot = T)#OK!

#Try removing interaction
LRS_HF_m2 <- lm(Totrep ~ Treatment + Strain, data = totalrep_HF)
summary(LRS_HF_m2)#SKN-1 have (significantly) lower LRS now in hot first
Anova(LRS_HF_m2, type = 'II')
#simulate residuals
LRS_HF_sim2 <- simulateResiduals(LRS_HF_m2, plot = T)#OK!


#Dataframe into long form to start fresh
thermo_h_long <- thermo_HF_rep %>% 
  pivot_longer(
    cols = c(`D1`,`D2`, `D3`, `D4`, `D5`,`D6`, `D7`,`D8`,`D9`), 
    names_to = "Day",
    values_to = "value")

#create a level order for 10 days
level_order <- c('D1', 'D2', 'D3', 'D4', 'D5','D6','D7','D8','D9')

#reorder levels so 10 is last
thermo_h_long$Day <- factor(thermo_h_long$Day, levels = level_order)


#spread data back out
thermo_h_wide <- spread(thermo_h_long, Day, value)
str(thermo_h_wide)

thermo_h_wide <- thermo_h_wide %>%
  dplyr::select(-c(Matricide))

thermo_h_wide <- na.omit(thermo_h_wide)

#calculate Lambda for each individual
L <- matrix(nrow = nrow(thermo_h_wide), ncol = 4)

for (i in 1:nrow(thermo_h_wide)){
  Les <- matrix(0, ncol = 11, nrow = 11)
  diag(Les[-1,]) <- rep(1, 10) # add the 1s for survival probability diagonally
  Fert <- c(0,0, as.numeric(as.vector(thermo_h_wide[i,][4:12]))) #the columns in data that has the reproductive counts
  Fert[is.na(Fert)] <- 0 #makes all NAs into 0s
  Les[1,] <- c(Fert)
  class(Les) <- "leslie.matrix"
  Lambda <- popbio::eigen.analysis(Les)$lambda1
  L[i, 1:4] <- c(paste0(thermo_h_wide$ID[i]), paste0(thermo_h_wide$Strain[i]), paste0(thermo_h_wide$Treatment[i]), Lambda)
  
}

#rename columns
colnames(L)<-c("ID", "Strain", "Treatment", "Lambda")

#make L a data frame
thermo_h_lam<-as.data.frame(L)

#make sure it's numeric
thermo_h_lam$Lambda<-as.numeric(as.character(thermo_h_lam$Lambda))
thermo_h_lam$Treatment <- as.factor(thermo_h_lam$Treatment)


#lambda analysis
hist(thermo_h_lam$Lambda)#a bit skewed, but OK
thermo_h_lam_m1 <- lm(Lambda ~ Strain * Treatment, data = thermo_h_lam)
summary(thermo_h_lam_m1)
sim_lam_m1 <- simulateResiduals(thermo_c_lam_m2, plot = T)#ok
Anova(thermo_h_lam_m1, type = 'II')


```

## Lifespan
```{r}
#Combine Strain and Treatment into one variable for plotting
thermo_HF_LS <- thermo_HF_LS %>%
  unite(Str_tr, c('Strain', 'Treatment.ID'), remove = F)

#Make sure it's a factor
thermo_HF_LS$Str_tr <- as.factor(thermo_HF_LS$Str_tr)

#Set N2 Ev as reference level
thermo_HF_LS$Str_tr <- relevel(thermo_HF_LS$Str_tr, ref = 'N2_ev')

#FIlter out unnatural deaths (Loss, walling, explosions)
thermo_HF_LS <- thermo_HF_LS %>%
  filter(thermo_HF_LS$Cause != 'L' & thermo_HF_LS$Cause !='W' & thermo_HF_LS$Cause != 'E')

#name factor levels of Str_tr
thermo_HF_LS$Str_tr <- factor(thermo_HF_LS$Str_tr, levels = c("N2_ev", "N2_daf", "SKN-1_ev", "SKN-1_daf"))

#Check what Age is classed as
class(thermo_HF_LS$Age)

#Change to numeric
thermo_HF_LS$Age <- as.numeric(thermo_HF_LS$Age)

#Create survival object
survHF<-survfit(Surv(Age,Event)~Str_tr,data=thermo_HF_LS)

#Create a binary variable for matricides (1 = matricide, 0 = no matricide) to analyse with logistic regression
thermo_HF_LS$Matricide <- as.numeric(thermo_HF_LS$Cause == 'M')

#binomial glmer for matricide occurrence (logistic regression)
mat_mod <- glmer(Matricide ~ Str_tr + (1|Plate.ID), data = thermo_HF_LS, family = 'binomial')
summary(mat_mod)

#Exclude matricides
thermo_HF_LSNM <- thermo_HF_LS %>%
  filter(thermo_HF_LS$Cause != 'M')

#Sometimes NAs are produced creating an additional non-existent factor level - omit NAs first
thermo_HF_LS <- na.omit(thermo_HF_LS)
thermo_HF_LSNM <- na.omit(thermo_HF_LSNM)

#Drop non-existent factor level - or cox models get confused
thermo_HF_LS$Str_tr <- droplevels(thermo_HF_LS$Str_tr)
levels(thermo_HF_LS$Str_tr)
thermo_HF_LSNM$Str_tr <- droplevels(thermo_HF_LSNM$Str_tr)

#Mixed effects cox proportional hazards model to test effect of treatment on survival probability
HF_cox <- coxme(Surv(Age, Event) ~ Strain * Treatment.ID + (1|Plate.ID), data = thermo_HF_LS)
summary(HF_cox)
Anova(HF_cox, type = 'III')

#Test model 
cox.zph(HF_cox)#check with EH analysis

#Mixed effects cox proportional hazards model to test effect of treatment on survival prob
HF_NM_cox <- coxme(Surv(Age, Event) ~ Strain * Treatment.ID + (1|Plate.ID), data = thermo_HF_LSNM)
summary(HF_NM_cox)
Anova(HF_NM_cox, type = 'III')
#test model
plot(cox.zph(HF_NM_cox))
cox.zph(HF_NM_cox)#compare to an EH analysis

#Check with EH analysis

#Matricides included
#Change the format of the Date columns to ensure that R recognises them as actual dates
thermo_HF_LS_EH <- thermo_HF_LS %>%
  mutate(
    D1 = as.Date(paste0(D1, "-2025"), format = "%d-%b-%Y"),
    Death.Date = as.Date(paste0(Death.Date, "-2025"), format = "%d-%b-%Y")
  )

#Create an individual ID number for every worm
thermo_HF_LS_EH <- thermo_HF_LS_EH %>%
  mutate(ID = row_number())

#Create a new column status which contains '0' for any day the worm is alive and '1' for the date that it dies
EH_thermo_HF_LS <- thermo_HF_LS_EH %>%
  filter(!is.na(D1), !is.na(Death.Date)) %>%
  rowwise() %>%
  mutate(Date = list(seq(D1, Death.Date, by = "1 day"))) %>%
  unnest(Date) %>%
  ungroup() %>%
  group_by(ID) %>%
  mutate(
    Day = row_number(),    # day counter per individual
    status = if_else(Date == Death.Date, 1, 0)
  ) %>%
  ungroup() %>%
  dplyr::select(c(ID, Strain, Treatment.ID, Plate.ID, Worm, Cause, Date, Day, status, Age, Event))

#remove causes of death we want censored
EH_thermo_HF_LS <- EH_thermo_HF_LS %>%
  filter(Cause != 'W' & Cause != 'E' & Cause != 'L')

#make sure this all worked!
head(EH_thermo_HF_LS)

#run binomial glmer - include ID for repeated measures as random factor and Day to account for time on survival
EH_m1 <- glmmTMB(status ~ Strain * Treatment.ID + (1 | ID) + (1 | Day), family = "binomial",
    data = EH_thermo_HF_LS)

summary(EH_m1)#these results are the same as cox model - will use the cox model for consistency



#Matricides censored
#Change the format of the Date columns to ensure that R recognises them as actual dates
thermo_HF_LS_EHNM <- thermo_HF_LSNM %>%
  mutate(
    D1 = as.Date(paste0(D1, "-2025"), format = "%d-%b-%Y"),
    Death.Date = as.Date(paste0(Death.Date, "-2025"), format = "%d-%b-%Y")
  )

#Create an individual ID number for every worm
thermo_HF_LS_EHNM <- thermo_HF_LS_EHNM %>%
  mutate(ID = row_number())

#Create a new column status which contains '0' for any day the worm is alive and '1' for the date that it dies
EH_thermo_HF_LSNM <- thermo_HF_LS_EHNM %>%
  filter(!is.na(D1), !is.na(Death.Date)) %>%
  rowwise() %>%
  mutate(Date = list(seq(D1, Death.Date, by = "1 day"))) %>%
  unnest(Date) %>%
  ungroup() %>%
  group_by(ID) %>%
  mutate(
    Day = row_number(),    # day counter per individual
    status = if_else(Date == Death.Date, 1, 0)
  ) %>%
  ungroup() %>%
  dplyr::select(c(ID, Strain, Treatment.ID, Plate.ID, Worm, Cause, Date, Day, status, Age, Event))

#remove causes of death we want censored
EH_thermo_HF_LSNM <- EH_thermo_HF_LSNM %>%
  filter(Cause != 'W' & Cause != 'E' & Cause != 'L')

#make sure this all worked!
head(EH_thermo_HF_LSNM)

#run binomial glmer - include ID for repeated measures as random factor and Day to account for time on survival
EH_m1_NM <- glmmTMB(status ~ Strain * Treatment.ID + (1 | ID) + (1 | Day), family = "binomial",
    data = EH_thermo_HF_LSNM)

summary(EH_m1_NM)#these results are the same as cox model - will use the cox model for consistency

```


# SKN-1 in 15C
Lifespan assays in SKN-1 mutants and N2s with daf-2 or ev (control) RNAi treatments in 15C, including a treatment with a single 24-hour cycle of 25C at ~ L2 stage
##Lifespan
```{r}

#Include only 15C data
SKN_15 <- SKN_15_25_LS %>%
  filter(Temperature != '25' & Temperature != '25F')

#Unite strain and treatment for plotting
SKN_15 <- SKN_15 %>%
  unite(Str_tr, 'Strain','Treatment.ID', remove = F)

#Unite strain, treatment and temperature for plotting
SKN_15 <- SKN_15 %>%
  unite(Str_tr_temp, 'Str_tr', 'Temperature', remove = F)

#Make sure it's a factor
SKN_15$Str_tr_temp <- as.factor(SKN_15$Str_tr_temp)

#Set reference level to N2 Ev
SKN_15$Str_tr_temp <- relevel(SKN_15$Str_tr_temp, ref = 'N2_ev_15')

#Name levels
SKN_15$Str_tr_temp <- factor(SKN_15$Str_tr_temp, levels = c('N2_ev_15','N2_daf_15','SKN_ev_15','SKN_daf_15','SKN_daf_15F'))

#Exclude matricides
SKN_15_NM <- SKN_15 %>%
  filter(Cause != 'M')

#Analyse two strains individually due to additionally treatment of thermocycling during development in SKN-1 only
#FIlter only SKNs 
SKN_only_15 <- SKN_15 %>%
  filter(Strain == 'SKN')

str(SKN_only_15)

#Check levels of Strain
SKN_only_15$Strain <- as.factor(SKN_only_15$Strain)
levels(SKN_only_15$Strain)

#Check levels of temperature
levels(SKN_only_15$Temperature)

#Need to drop levels - still thinks it has 15C treatments 
SKN_only_15$Temperature <- droplevels(SKN_only_15$Temperature)
levels(SKN_only_15$Temperature)

#Check levels of Treatment
SKN_only_15$Treatment.ID <- as.factor(SKN_only_15$Treatment.ID)
levels(SKN_only_15$Treatment.ID)
SKN_only_15$Treatment.ID <- droplevels(SKN_only_15$Treatment.ID)

#Combine Treatment and temperature because of imbalance
SKN_only_15 <- SKN_only_15 %>%
  unite(Tr_temp, 'Treatment.ID','Temperature', remove = F)

#set ev as reference level
SKN_only_15$Tr_temp <- as.factor(SKN_only_15$Tr_temp)
SKN_only_15$Tr_temp <- relevel(SKN_only_15$Tr_temp, ref = 'ev_15')

#Coxme model with combined effect of treatment and temperature to account for developmental thermocycle treatment
cox_SKN_15_only <- coxme(Surv(Age, Event) ~ Tr_temp + (1|Plate.ID), data = SKN_only_15)
summary(cox_SKN_15_only)
Anova(cox_SKN_15_only, type = 'II')

#test model
cox.zph(cox_SKN_15_only)# OK! p = 0.18

#Matricides censored
SKN_only_15_NM <- SKN_only_15 %>%
  filter(Cause !='M')

#Coxme model with combined effect of treatment and temperature to account for developmental thermocycle treatment with matricides censored
cox_SKN_15_only_NM <- coxme(Surv(Age, Event) ~ Tr_temp + (1|Plate.ID), data = SKN_only_15_NM)
summary(cox_SKN_15_only_NM)
Anova(cox_SKN_15_only_NM, type = 'II')
#test model
cox.zph(cox_SKN_15_only_NM)# global p = 0.41, OK!


#Filter out N2s
N2_only_15 <- SKN_15 %>%
  filter(Strain == 'N2')

#Set ev as reference level
N2_only_15$Treatment.ID <- as.factor(N2_only_15$Treatment.ID)
N2_only_15$Treatment.ID <- relevel(N2_only_15$Treatment.ID, ref = 'ev')

#Coxme model with treatment as factor
cox_N2_15_only <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = N2_only_15)
summary(cox_N2_15_only)
Anova(cox_N2_15_only, type = 'II')

#test model
cox.zph(cox_N2_15_only)#OK p = 0.79

#Censor matricides
N2_only_15_NM <- N2_only_15 %>%
  filter(Cause !='M')

#coxme model with treatment as factor with matricides censored
cox_N2_15_only_NM <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = N2_only_15_NM)
summary(cox_N2_15_only_NM)
Anova(cox_N2_15_only_NM, type = 'II')
#test model
cox.zph(cox_N2_15_only_NM)#p = 0.87 OK


```

# SKN-1 in 25C
Lifespan assays in SKN-1 mutants and N2s with daf-2 or ev (control) RNAi treatments in 25C, including a treatment with a single 24-hour cycle of 15C at ~ L2 stage
## Lifespan
```{r}

#Temperature as a factor
SKN_15_25_LS$Temperature <- as.factor(SKN_15_25_LS$Temperature)

#Get only 25C data
SKN_25 <- SKN_15_25_LS %>%
  filter(Temperature != '15' & Temperature != '15F')

#Combine Strain and treatment for plotting
SKN_25 <- SKN_25 %>%
  unite(Str_tr, 'Strain','Treatment.ID', remove = F)

#Also combine temperature with strain and treatment
SKN_25 <- SKN_25 %>%
  unite(Str_tr_temp, 'Str_tr', 'Temperature', remove = F)

#Rename factor levels
SKN_25$Str_tr_temp <- factor(SKN_25$Str_tr_temp, levels = c('N2_ev_25','N2_daf_25','SKN_ev_25','SKN_daf_25','SKN_daf_25F'))

#Ensure it's a factor
SKN_25$Str_tr_temp <- as.factor(SKN_25$Str_tr_temp)

#Set ev N2 with 25C only as reference level
SKN_25$Str_tr_temp <- relevel(SKN_25$Str_tr_temp, ref = 'N2_ev_25')


#Exclude matricides
SKN_25_NM <- SKN_25 %>%
  filter(Cause != 'M')

#Make sure Str-tr_temp is a factor
SKN_25_NM$Str_tr_temp <- as.factor(SKN_25_NM$Str_tr_temp)

#Reset baseline to ev N2 in 25 only
SKN_25_NM$Str_tr_temp <- relevel(SKN_25_NM$Str_tr_temp, ref = 'N2_ev_25')


#Analyse within strains becaues of additional treatment only in SKN-1

#FIlter only SKNs 
SKN_only_25 <- SKN_25 %>%
  filter(Strain == 'SKN')

str(SKN_only_25)

#Check levels of Strain
SKN_only_25$Strain <- as.factor(SKN_only_25$Strain)
levels(SKN_only_25$Strain)

#Check levels of temperature
levels(SKN_only_25$Temperature)

#Need to drop levels - still thinks it has 15C treatments 
SKN_only_25$Temperature <- droplevels(SKN_only_25$Temperature)
levels(SKN_only_25$Temperature)

#Check levels of Treatment
SKN_only_25$Treatment.ID <- as.factor(SKN_only_25$Treatment.ID)
levels(SKN_only_25$Treatment.ID)
SKN_only_25$Treatment.ID <- droplevels(SKN_only_25$Treatment.ID)

#Combine Treatment and temperature because of imbalance
SKN_only_25 <- SKN_only_25 %>%
  unite(Tr_temp, 'Treatment.ID','Temperature', remove = F)

#set ev as reference level
SKN_only_25$Tr_temp <- as.factor(SKN_only_25$Tr_temp)
SKN_only_25$Tr_temp <- relevel(SKN_only_25$Tr_temp, ref = 'ev_25')

#Coxme model with combined effect of treatment and temperature to account for developmental thermocycle treatment
cox_SKN_25_only <- coxme(Surv(Age, Event) ~ Tr_temp + (1|Plate.ID), data = SKN_only_25)
summary(cox_SKN_25_only)
Anova(cox_SKN_25_only, type = 'II')
#test model
cox.zph(cox_SKN_25_only)# global p = 0.004 - check with EH analysis

#Matricides censored
SKN_only_25_NM <- SKN_only_25 %>%
  filter(Cause !='M')

#Coxme model with combined effect of treatment and temperature to account for developmental thermocycle treatment with matricides censored
cox_SKN_25_only_NM <- coxme(Surv(Age, Event) ~ Tr_temp + (1|Plate.ID), data = SKN_only_25_NM)
summary(cox_SKN_25_only_NM)
Anova(cox_SKN_25_only_NM, type = 'II')

#test model
cox.zph(cox_SKN_25_only_NM)# global p = 0.054 OK


#Filter out N2s
N2_only_25 <- SKN_25 %>%
  filter(Strain == 'N2')

#Set ev as reference level
N2_only_25$Treatment.ID <- as.factor(N2_only_25$Treatment.ID)
N2_only_25$Treatment.ID <- relevel(N2_only_25$Treatment.ID, ref = 'ev')

#Coxme model with treatment as factor
cox_N2_only <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = N2_only_25)
summary(cox_N2_only)
Anova(cox_N2_only, type = 'II')

#test model
cox.zph(cox_N2_only)#OK p>0.05

#Censor matricides
N2_only_25_NM <- N2_only_25 %>%
  filter(Cause !='M')

#coxme model with treatment as factor with matricides censored
cox_N2_only_NM <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = N2_only_25_NM)
summary(cox_N2_only_NM)
Anova(cox_N2_only_NM, type = 'II')

#test model
cox.zph(cox_N2_only_NM)#p > 0.05 OK

#Check with EH analysis

#Matricides included
#Change the format of the Date columns to actual dates and add ID
SKN_only_25_EH<- SKN_only_25 %>%
  mutate(
    D1 = as.Date(paste0(D1, "-2025"), format = "%d-%b-%Y"),
    Death.Date = as.Date(paste0(Death.Date, "-2025"), format = "%d-%b-%Y"))%>%
  mutate(ID = row_number())



#Create a new column status which contains '0' for any day the worm is alive and '1' for the date that it dies
EH_SKN_only_25 <- SKN_only_25_EH %>%
  filter(!is.na(D1), !is.na(Death.Date)) %>%
  rowwise() %>%
  mutate(Date = list(seq(D1, Death.Date, by = "1 day"))) %>%
  unnest(Date) %>%
  ungroup() %>%
  group_by(ID) %>%
  mutate(
    Day = row_number(),    # day counter per individual
    status = if_else(Date == Death.Date, 1, 0)
  ) %>%
  ungroup() %>%
  dplyr::select(c(ID, Strain, Tr_temp,Treatment.ID, Plate.ID, Worm, Cause, Date, Day, status, Age, Event))

#remove causes of death we want censored
EH_SKN_only_25 <- EH_SKN_only_25 %>%
  filter(Cause != 'W' & Cause != 'E' & Cause != 'L')

#make sure this all worked!
head(EH_SKN_only_25)

#run binomial glmer - include ID for repeated measures as random factor and Day to account for time on survival
SKN_25_EH_m1 <- glmmTMB(status ~ Tr_temp + (1 | ID) + (1 | Day), family = "binomial",
    data = EH_SKN_only_25)

summary(SKN_25_EH_m1)#these results are the same as cox model - will use the cox model for consistency
summary(cox_SKN_25_only)


```