---
title: "plots"
author: "Sara Irish"
date: "`r Sys.Date()`"
output: html_document
---

#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages({
  library(glmmTMB)
  library(car)
  library(popbio)
  library(ggplot2)
  library(ggbeeswarm)
  library(shades)
  library(RColorBrewer)
  library(DHARMa)
  library(Rmisc)
  library(dabestr)
  library(magrittr)
  library(tidyr)
  library(dplyr)
  library(ggeffects)
  library(lme4)
  library(optimx)
  library(nloptr)
  library(dfoptim)
  library(data.table)
  library(pscl)
  library(ggThemeAssist)
  library(DHARMa)
  library(emmeans)
  library(ggpubr)
  library(patchwork)
  library(lubridate)
  library(survival)
  library(coxme) #for cox model
  library(ggplot2)
  library(purrr) #for forestplot
  library(ggpubr) #for forestplot
  library(survminer) #for forestplot
  library(AICcmodavg)
  library(permutes)
  library(performance)
  library(fields)
  library(grid)
  library(ggplotify)
  library(cowplot)
  library(eha)
  library(multcomp)
  library(usethis)
  library(forcats)
})

getwd()
setwd("C:/Users/irish/OneDrive/Documents/R/SKN-1_IIS/SKN-1_IIS")
#24-hour 15 to 25C thermocycling experiments, HF = 25C first, CF = 15C first, LS = lifespan, rep = reproduction
thermo_HF_rep <- read.csv('data/Thermo_H_rep.csv')
thermo_HF_LS <- read.csv('data/Thermo_H_LS.csv')
thermo_CF_rep <- read.csv('data/Thermo_C_rep.csv')
thermo_CF_LS <- read.csv('data/Thermo_C_LS.csv') 

#SKN-1 mutants with and without daf-2 RNAi KD in 20C - LS = lifespan, rep = reproduction
SKN_20C_rep <- read.csv('data/SKN1_20C_repro.csv')
SKN_20C_LS <- read.csv('data/SKN1_20C_LS.csv')

#Lifespan experiment SKN-1 mutants with and without daf-2 RNAi KD in 15 and 25C, with additional treatment of daf-2 RNAi KD SKN-1 mutants starting with 24 hrs of opposite temperature to test effects of temperature on early development of SKN-1 mutants
SKN_15_25_LS <- read.csv('data/SKN_N2_LS_15_25.csv')

#Create colour palette
palette <- c('#E64B35FF','#4DBBD5FF', '#00A087FF', '#3C5488FF')

#plot labels for survival plots
N2_daf_label <- expression("N2: "*italic("daf-2"))
skn_daf_label <- expression(italic("skn-1: daf-2"))
skn_ev_label <- expression(italic("skn-1: ")*'ev')
skn_cycle_label <- expression(italic('skn-1: daf-2')*' + cycle')

lam_lab <- expression(Delta*' Individual fitness (vs N2 ev)')
lrs_lab <- expression(Delta*' LRS (vs N2 ev)')
skn_lab <- expression(italic('skn-1'))
daf_label <- expression(italic('daf-2'))
daf_cyc_forest_lab <- expression(
  atop(
    italic('daf-2'),
    '+cycle'))


#plot lables for forest plots
N2_daf_forest_lab <- expression(
  atop(
    'N2:',
    italic('daf-2')*' - ev'))

skn_daf_forest_lab <- expression(
  atop(
    italic('skn-1:'),
    italic('daf-2')*' - ev'))


GAM_skn_daf_forest_lab <- expression(
  atop(
    italic('skn-1:'),
    italic('daf-2')*' / ev'))

skn_daf_cyc_forest_lab <- expression(
  atop(
    italic('skn-1:'),
    italic('daf-2')*'+cycle - ev'))



#function to get contrasts from linear and Gamma models
get_contrasts <- function(model, ref = 1, type = NULL) {
  emm <- emmeans(model, ~ Treatment, type = type)
  contr <- contrast(emm, method = "trt.vs.ctrl", ref = ref)
  as.data.frame(confint(contr))
}

#function to build forest plot for just daf-2/ev comparisons within Skn-1
forest_plot <- function(df, xlab, ref_line) {
  ggplot(df, aes(x = estimate, y = contrast)) +
    geom_point(size = 3, colour = "#4DBBD5FF") +
    geom_errorbarh(
      aes(xmin = lower.CL, xmax = upper.CL),
      height = 0.2,
      colour = "#4DBBD5FF"
    ) +
    geom_vline(
      xintercept = ref_line,
      linetype = "dashed",
      colour = "#E64B35FF",
      linewidth = 1
    ) +
    labs(x = xlab, y = NULL) +
    theme_minimal(base_size = 12) +
    theme(
      legend.position = "none",
      axis.text = element_text(size = 12),
      axis.title = element_text(size = 12),
      axis.line.x = element_line(color = "black", size = 0.5),
      axis.line.y = element_line(color = "black", size = 0.5))
}

#forest plot function to compare across N2/skn-1 and daf-2/ev combos
forest_emmeans <- function(model,
                           strain = "Strain",
                           treatment = "Treatment",
                           ref_strain = "N2",
                           ref_treatment = "ev",
                           level = 0.95,
                           contrast_order = c(
                             "N2 daf - N2 ev",
                             "(SKN-1 ev) - N2 ev",
                             "(SKN-1 daf) - N2 ev"
                           ),
                           xlab = NULL) {
  
  # Estimated marginal means
  emm <- emmeans::emmeans(
    model,
    as.formula(paste("~", strain, "*", treatment))
  )
  
  # Index of reference
  ref_idx <- which(
    emm@grid[[strain]] == ref_strain &
    emm@grid[[treatment]] == ref_treatment
  )
  
  # Contrasts vs reference
  ctr <- emmeans::contrast(emm, "trt.vs.ctrl", ref = ref_idx)
  
  # Confidence intervals
  df <- as.data.frame(confint(ctr, level = level))
  
  # Set contrast order if specified
  if(!is.null(contrast_order)){
    df$contrast <- factor(df$contrast, levels = contrast_order)
  }
  
  # Default x-axis label if none provided
  if(is.null(xlab)){
    xlab <- paste0("Effect size (vs ", ref_strain, " ", ref_treatment, ")")
  }
  
  # Plot
  ggplot2::ggplot(df, ggplot2::aes(x = estimate, y = contrast, color = contrast)) +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "#E64B35FF") +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_errorbarh(
      ggplot2::aes(xmin = lower.CL, xmax = upper.CL),
      height = 0.2
    ) +
    ggplot2::labs(
      x = xlab,
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 12)+
    ggplot2::theme(legend.position = 'none',
        axis.text = element_text(size = 12),
        axis.line.x = element_line(color = "black", size = 0.5),
      axis.line.y = element_line(color = "black", size = 0.5))
  
}



#function for coxme forest plots for thermocycling
forest_coxme <- function(model,
                          strain = "Strain",
                          treatment = "Treatment.ID",
                          ref_strain = "N2",
                          ref_treatment = "ev",
                          level = 0.95,
                          xlab = "Log Hazard Ratio") {
  
  # Fixed contrast order
  contrast_order <- c(
    "N2 daf - N2 ev",
    "(SKN-1 ev) - N2 ev",
    "(SKN-1 daf) - N2 ev"
  )
  
  # Estimated marginal means
  emm <- emmeans::emmeans(model, as.formula(paste("~", strain, "*", treatment)))
  
  # Reference index
  ref_idx <- which(
    emm@grid[[strain]] == ref_strain &
    emm@grid[[treatment]] == ref_treatment
  )
  
  # Contrasts vs baseline
  ctr <- emmeans::contrast(emm, "trt.vs.ctrl", ref = ref_idx)
  
  # Convert to data frame
  df <- as.data.frame(ctr)
  
  # Manually compute confidence intervals
  z <- qnorm(1 - (1 - level)/2)
  df$lower.CL <- df$estimate - z * df$SE
  df$upper.CL <- df$estimate + z * df$SE
  
  # Apply fixed contrast order
  df$contrast <- factor(df$contrast, levels = contrast_order)
  
  # Plot
  ggplot2::ggplot(df, ggplot2::aes(x = estimate, y = contrast, color = contrast)) +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "#E64B35FF") +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_errorbarh(
      ggplot2::aes(xmin = lower.CL, xmax = upper.CL),
      height = 0.2
    ) +
    ggplot2::labs(
      x = xlab,
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 13) +
    ggplot2::theme(
      axis.line = ggplot2::element_blank(),
      axis.line.x = ggplot2::element_line(color = "black", size = 0.5),
      axis.line.y = ggplot2::element_line(color = "black", size = 0.5),
      legend.position = 'none',
      axis.text = element_text(size = 12)
    )
}


#function for coxme forest plots for different temperature treatments + short thermocycle 
forest_coxme <- function(model,
                          strain = "Strain",
                          treatment = "Treatment.ID",
                          ref_strain = "N2",
                          ref_treatment = "ev",
                          level = 0.95,
                          xlab = "Log Hazard Ratio") {
  
  # Fixed contrast order
  contrast_order <- c(
    "N2 daf - N2 ev",
    "(SKN-1 ev) - N2 ev",
    "(SKN-1 daf) - N2 ev"
  )
  
  # Estimated marginal means
  emm <- emmeans::emmeans(model, as.formula(paste("~", strain, "*", treatment)))
  
  # Reference index
  ref_idx <- which(
    emm@grid[[strain]] == ref_strain &
    emm@grid[[treatment]] == ref_treatment
  )
  
  # Contrasts vs baseline
  ctr <- emmeans::contrast(emm, "trt.vs.ctrl", ref = ref_idx)
  
  # Convert to data frame
  df <- as.data.frame(ctr)
  
  # Manually compute confidence intervals
  z <- qnorm(1 - (1 - level)/2)
  df$lower.CL <- df$estimate - z * df$SE
  df$upper.CL <- df$estimate + z * df$SE
  
  # Apply fixed contrast order
  df$contrast <- factor(df$contrast, levels = contrast_order)
  
  # Plot
  ggplot2::ggplot(df, ggplot2::aes(x = estimate, y = contrast, color = contrast)) +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "#E64B35FF") +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_errorbarh(
      ggplot2::aes(xmin = lower.CL, xmax = upper.CL),
      height = 0.2
    ) +
    ggplot2::labs(
      x = xlab,
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 13) +
    ggplot2::theme(
      axis.line = ggplot2::element_blank(),
      axis.line.x = ggplot2::element_line(color = "black", size = 0.5),
      axis.line.y = ggplot2::element_line(color = "black", size = 0.5),
      legend.position = 'none',
      axis.text = element_text(size = 12)
    )
}


forest_cox_strain <- function(model,
                              treatment = "Treatment.ID",
                              ref_treatment = NULL,
                              level = 0.95,
                              xlab = "Log Hazard Ratio",
                              treatment_labels = NULL,
                              vline_color = "grey50") {
  
  # Estimated marginal means
  emm <- emmeans::emmeans(model, as.formula(paste("~", treatment)))
  
  # Reference index
  if(!is.null(ref_treatment)){
    ref_idx <- which(emm@grid[[treatment]] == ref_treatment)
    ctr <- emmeans::contrast(emm, "trt.vs.ctrl", ref = ref_idx)
  } else {
    ctr <- emmeans::contrast(emm, "pairwise")
  }
  
  # Convert to data frame
  df <- as.data.frame(ctr)
  
  # Compute CI manually
  z <- qnorm(1 - (1 - level)/2)
  df$lower.CL <- df$estimate - z * df$SE
  df$upper.CL <- df$estimate + z * df$SE
  
  # Create a clean label column for plotting
  df$Treatment <- gsub(paste0(" - .*"), "", df$contrast)
  
  # Apply optional treatment label order
  if(!is.null(treatment_labels)){
    df$Treatment <- factor(df$Treatment, levels = treatment_labels)
  } else {
    df$Treatment <- factor(df$Treatment)
  }
  
  # Plot
  ggplot2::ggplot(df, ggplot2::aes(x = estimate, y = Treatment, color = Treatment)) +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = vline_color) +
    ggplot2::geom_point(size = 3) +
    ggplot2::geom_errorbarh(
      ggplot2::aes(xmin = lower.CL, xmax = upper.CL),
      height = 0.2
    ) +
    ggplot2::labs(
      x = xlab,
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 13) +
    ggplot2::theme(
      axis.line = ggplot2::element_blank(),
      axis.text = ggplot2::element_text(size = 12),
      axis.line.x = ggplot2::element_line(color = "black", size = 0.5),
      axis.line.y = ggplot2::element_line(color = "black", size = 0.5)
    )
}



```


#SKN-1 in 20C
Reproduction and lifespan assays in SKN-1 mutants with daf-2 or ev (control) RNAi treatments in 20C

##Reproduction
```{r}

#put data into long form
skn_long <- SKN_20C_rep %>% 
  pivot_longer(
    cols = c(`D1`, `D2`, `D3`, `D4`, `D5`), 
    names_to = "Day",
    values_to = "value")

summarySE(thermo_HF_LS, groupvars = c('Strain','Treatment.ID'), measurevar = 'Age')


#select only columns needed
skn_long <- skn_long %>%
  dplyr::select(c('ID','Treatment','Day','value'))

#rename treatments
skn_long$Treatment <- as.factor(skn_long$Treatment)
levels(skn_long$Treatment) <- list('ev' = 'ev', 'daf-2' = 'daf')


#get rid of NAs
skn_long <- na.omit(skn_long)

#age-specific reproduction plot for SKN-1 mutants with daf or ev
skn_20C_p <-ggplot(data=skn_long, aes(x=factor(Day), y=value, group=Treatment, color=Treatment))+
  geom_jitter(alpha = 0.2, position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.5))+
  stat_summary(fun.data="mean_cl_boot", geom="errorbar", size = 1.2, width=0.0, position = position_dodge(0.5)) +
  stat_summary(fun.data="mean_cl_boot", geom="point", size = 3, position = position_dodge(0.5)) +
  stat_summary(fun.data="mean_cl_boot", geom="line",  size=1.2, position = position_dodge(0.5)) +
  theme_classic()+
  labs(y="Offspring number", x="Day")+
  labs(col="")+
  theme(axis.title.y = element_text(size=14))+
  theme(axis.title.x = element_text(size=14))+
  theme(axis.text= element_text(size = 12))+
  theme(legend.text = element_text(size = 12))+
  scale_color_manual(values=c('#E64B35FF','#4DBBD5FF'), labels = c(skn_ev_label, skn_daf_label))+
  theme(legend.key.width = unit(0.5,"cm"))+
  coord_cartesian(ylim = c(0,225))+
  theme(legend.position = c(0.1, 0.9))

skn_20C_p

#Get Day into a numeric form
#Change day to a numeric variable for analysis
skn_long$Day <- as.factor(skn_long$Day)
levels(skn_long$Day) <- list('1' = 'D1', '2' = 'D2', '3' = 'D3', '4' = 'D4', '5' = 'D5')
skn_long$Day <- as.numeric(skn_long$Day)

#set ev as reference level
skn_long$Treatment <- as.factor(skn_long$Treatment)
skn_long$Treatment <- relevel(skn_long$Treatment, ref = 'ev')

##############
#### LRS
##############

#sum up total number of offspring produced over experiment per individual
totalrep_SKN<-na.omit(as.data.frame.table(tapply(skn_long$value,list(skn_long$Treatment, skn_long$ID),sum)))

#rename columns
names(totalrep_SKN)<-c("Treatment", "Replicate", "Totrep")


#Try a Gamma GLM
SKN_LRS_m2 <- glm(Totrep ~ Treatment, data = totalrep_SKN, family = Gamma(link = 'log'))

#get contrasts
lrs_df <- get_contrasts(SKN_LRS_m2, type = "response")


#change name of ratio variable to estimate, just to make plot function work
str(lrs_df)
lrs_df <- lrs_df %>%
  dplyr::rename(
    estimate = ratio
  )

#LRS forest plot 
skn_LRS_forest <- forest_plot(
  lrs_df,
  xlab = "LRS mean ratio (vs ev)",
  ref_line = 1
)

skn_LRS_forest <- skn_LRS_forest + scale_y_discrete(labels = c(GAM_skn_daf_forest_lab))

skn_LRS_forest


###########
#lambda#
##########

#### Lambda
#start fresh
skn_long <- SKN_20C_rep %>% 
  pivot_longer(
    cols = c(`D1`, `D2`, `D3`, `D4`, `D5`), 
    names_to = "Day",
    values_to = "value")

str(skn_long)
#keep only columns needed
skn_long <- skn_long %>%
  dplyr::select(c('ID', 'Strain','Treatment','Day','value'))
str(skn_long)


#spread data back out
skn_wide <- spread(skn_long, Day, value)


skn_wide <- na.omit(skn_wide)

#calculate Lambda for each individual
L <- matrix(nrow = nrow(skn_wide), ncol = 4)

for (i in 1:nrow(skn_wide)){
  Les <- matrix(0, ncol = 7, nrow = 7)
  diag(Les[-1,]) <- rep(1, 6) # add the 1s for survival probability diagonally
  Fert <- c(0,0, as.numeric(as.vector(skn_wide[i,][4:8]))) #the columns in data that has the reproductive counts
  Fert[is.na(Fert)] <- 0 #makes all NAs into 0s
  Les[1,] <- c(Fert)
  class(Les) <- "leslie.matrix"
  Lambda <- popbio::eigen.analysis(Les)$lambda1
  L[i, 1:4] <- c(paste0(skn_wide$ID[i]), paste0(skn_wide$Strain[i]), paste0(skn_wide$Treatment[i]), Lambda)
  
}

#rename columns
colnames(L)<-c("ID", "Strain", "Treatment", "Lambda")

#make L a data frame
skn_lam<-as.data.frame(L)

#make sure it's numeric
skn_lam$Lambda<-as.numeric(as.character(skn_lam$Lambda))
skn_lam$Treatment <- as.factor(skn_lam$Treatment)

skn_lam$Treatment <- revalue(skn_lam$Treatment, c('ev' = 'control', 'daf' = 'daf-2'))
skn_lam$Treatment <- relevel(skn_lam$Treatment, ref = 'control')

#analysis 
skn_lam_mod <- lm(Lambda ~ Treatment, data = skn_lam)
summary(skn_lam_mod)
Anova(skn_lam_mod, type = 'II')


#get means
skn_mean_lam <- summarySE(data = skn_lam, measurevar = 'Lambda', groupvars = c('Treatment'))
skn_mean_lam

#get contrasts for lam and LRS
lam_df <- get_contrasts(skn_lam_mod)
lrs_df <- get_contrasts(SKN_LRS_m2, type = "response")

#create lambda forest plot
skn_lam_forest <- forest_plot(
  lam_df,
  xlab = expression(Delta*' Individual fitness (vs ev)'),
  ref_line = 0
)
skn_lam_forest

#add italicised label
skn_lam_forest <- skn_lam_forest + scale_y_discrete(labels = c(skn_daf_forest_lab))

skn_lam_forest



```

## Lifespan
```{r}

#Get rid of NAs
SKN_20C_LS <- na.omit(SKN_20C_LS)

#Rename daf to daf-2
SKN_20C_LS$Treatment.ID <- revalue(SKN_20C_LS$Treatment.ID, c('ev' = 'ev', 'daf' = 'daf-2'))

#Make sure Treatment is a factor
SKN_20C_LS$Treatment.ID <- as.factor(SKN_20C_LS$Treatment.ID)

#Make ev the baseline level
SKN_20C_LS$Treatment.ID <- relevel(SKN_20C_LS$Treatment.ID, ref = 'ev')

#Check factor levels of treatment
levels(SKN_20C_LS$Treatment.ID)

#Get rid of empty levels
SKN_20C_LS$Treatment.ID <- droplevels(SKN_20C_LS$Treatment.ID)

#Plate.ID as a factor not a numerical
SKN_20C_LS$Plate.ID <- as.factor(SKN_20C_LS$Plate.ID)

#Create survival object for SKN-1 mutants at 20C
surv_SKN<-survfit(Surv(Age,Event)~Treatment.ID,data=SKN_20C_LS)

#Create survival curve from survival object
SKN_20C_LS_plot <-ggsurvplot(surv_SKN, ylab="Survival probability\n", data = SKN_20C_LS, size= 0.8, font.ylab= 14, font.xlab= 14, palette=c('#E64B35FF','#4DBBD5FF'), legend = c(0.8, 0.8), legend.title = "", legend.labs = c('ev','daf-2'), title = "", censor = FALSE, xlab = "Day", xlim=c(0,35), break.time.by = 5, position= position_dodge(0.9), font.tickslab = c(12), font.legend = c(12))

SKN_20C_LS_plot <- SKN_20C_LS_plot$plot + scale_colour_manual(values = c('#E64B35FF','#4DBBD5FF'), labels = c(skn_ev_label,skn_daf_label))
SKN_20C_LS_plot


#Mixed effects cox model for a effect of daf-2 on SKN-1 mutant lifespan
cox_SKN_20C <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = SKN_20C_LS)
summary(cox_SKN_20C)
Anova(cox_SKN_20C, type = 'II')

#test model
cox.zph(cox_SKN_20C)#OK


#Use emmeans because of negligible effect of plate.ID
SKN_emm <- emmeans(cox_SKN_20C, ~ Treatment.ID)  # EMMs of treatment 
SKN_emm

#Get contrasts
contrasts_SKN <- contrast(SKN_emm, method = "revpairwise")

#make a dataframe out of the contrast summary
df_SKN <- as.data.frame(summary(contrasts_SKN, infer = TRUE))

# Rename columns
names(df_SKN)[names(df_SKN) == "estimate"] <- "logHR"
df_SKN$HR <- exp(df_SKN$logHR)
df_SKN$lower.HR <- exp(df_SKN$asymp.LCL)
df_SKN$upper.HR <- exp(df_SKN$asymp.UCL)

#fix labels
df_SKN$Comparison <- df_SKN$contrast

#Create forest plot for differences of treatments within strains
SKN_contrasts <- ggplot(df_SKN, aes(x = Comparison, y = logHR)) +
  geom_point(size = 4, color = '#4DBBD5FF') +
  geom_hline(yintercept = 0, linetype = "dashed", size = 1.2, color = '#E64B35FF', linewidth = 0.8) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.2, size = 1.1,color ='#4DBBD5FF')+
  coord_flip() +
  scale_x_discrete(labels = c(skn_daf_forest_lab))+
  labs(
    title = "",
    x = "",
    y = "Log hazard ratio (HR)"
  ) +
  theme_minimal(base_size = 12)+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        legend.position = 'none')

SKN_contrasts

#Exclude matricides
SKN_20C_LS_NM <- SKN_20C_LS %>%
  filter(Cause != 'M')


#Create survival object
surv_SKN_NM<-survfit(Surv(Age,Event)~Treatment.ID,data=SKN_20C_LS_NM)

#Survival plot
SKN_NM_LS_plot <-ggsurvplot(surv_SKN_NM, ylab="Survival probability", data = SKN_20C_LS_NM, size= 0.8, font.ylab= 14, font.xlab= 14, palette=c('#E64B35FF','#4DBBD5FF'), legend = c(0.8, 0.8), legend.title = "", legend.labs = c('ev','daf-2'), title = "", censor = FALSE, xlab = "Day", xlim=c(0,35), break.time.by = 5, position= position_dodge(0.9), font.tickslab = c(12), font.legend = c(12))

SKN_NM_LS_plot


SKN_NM_LS_plot <- SKN_NM_LS_plot$plot + scale_colour_manual(values = c('#E64B35FF','#4DBBD5FF'), labels = c(skn_ev_label,skn_daf_label))
SKN_NM_LS_plot

#Cox mixed effects model
cox_SKN_20C_NM <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = SKN_20C_LS_NM)
summary(cox_SKN_20C_NM)
Anova(cox_SKN_20C_NM, type = 'II')
#Test model
cox.zph(cox_SKN_20C_NM)#OK

#Use emmeans because of negligible effect of plate.ID
SKN_NM_emm <- emmeans(cox_SKN_20C_NM, ~ Treatment.ID)  # EMMs of treatment 
SKN_NM_emm

#Get contrasts
contrasts_SKN_NM <- contrast(SKN_NM_emm, method = "revpairwise")

#make a dataframe out of the contrast summary
df_SKN_NM <- as.data.frame(summary(contrasts_SKN_NM, infer = TRUE))

# Rename columns
names(df_SKN_NM)[names(df_SKN_NM) == "estimate"] <- "logHR"
df_SKN_NM$HR <- exp(df_SKN_NM$logHR)
df_SKN_NM$lower.HR <- exp(df_SKN_NM$asymp.LCL)
df_SKN_NM$upper.HR <- exp(df_SKN_NM$asymp.UCL)

#fix labels
df_SKN_NM$Comparison <- df_SKN_NM$contrast

#Create forest plot for differences of treatments within strains
SKN_NM_contrasts <- ggplot(df_SKN_NM, aes(x = Comparison, y = logHR)) +
  geom_point(size = 4, color = '#4DBBD5FF') +
  geom_hline(yintercept = 0, linetype = "dashed", size = 1.2, color = '#E64B35FF', linewidth = 0.8) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.2, size = 1.1,color ='#4DBBD5FF')+
  coord_flip() +
  scale_x_discrete(labels = c(skn_daf_forest_lab))+
  labs(
    title = "",
    x = "",
    y = "Log hazard ratio (HR)"
  ) +
  theme_minimal(base_size = 12)+
  theme(axis.text = element_text(size = 13),
        axis.title = element_text(size = 14),
        legend.position = 'none')

SKN_NM_contrasts

#need to run previous code chunk to combine all plots

#Group all plots for 20C skn-1 data
SKN_rep_LRSLAM <- ggarrange(skn_LRS_forest, skn_lam_forest, ncol = 2, labels = c('C','D'))
SKN_rep_LRSLAM

SKN_20_all <- ggarrange(SKN_NM_LS_plot, skn_20C_p, SKN_rep_LRSLAM, nrow = 3, heights = c(1.5, 1.5, 1), labels = c('A', 'B', ''), common.legend = T)
SKN_20_all


#ggsave(plot = SKN_20_all, "SKN_20C_all.png", bg = 'white',width = 7, height = 9, dpi = 300)

#Matricides included for suppmat
SKN_20C_LS_all <- ggarrange(SKN_20C_LS_plot, SKN_contrasts+theme(plot.margin = margin(t = 5, r = 45, b = 5, l = 45, unit = "pt")), nrow = 2, heights = c(2,1), common.legend = T, labels = c('A','B'))
SKN_20C_LS_all
#ggsave("SKN_20C_LS_all.png", plot = SKN_20C_LS_all, bg = 'white',width = 6, height = 6, dpi = 300, units = "in")


```


#15C first Thermocycling

Reproduction and lifespan assays following 24 hour thermo-cycling between 15 and 25C with 15C as starting temperature across four treatments: SKN-1 mutant + daf-2 RNAi, SKN-1 mutant + ev (control), N2 + daf-2 RNAi, N2 + ev

## Reproduction plots
```{r}

#Dataframe into long form
thermo_c_long <- thermo_CF_rep %>% 
  pivot_longer(
    cols = c(`D1`,`D2`, `D3`, `D4`, `D5`,`D6`, `D7`,`D8`,`D9`), 
    names_to = "Day",
    values_to = "value")


#Remove unneeded columns
thermo_c_long <- thermo_c_long %>%
  dplyr::select(c('ID','Strain', 'Treatment','Day','value'))

#Remove NAs
thermo_c_long <- na.omit(thermo_c_long)

#Combine strain and treatment into one variable for plotting
thermo_c_long <- thermo_c_long %>%
  unite(Str_tr, c('Strain', 'Treatment'), remove = F)

#Set levels of StrainxTreatment
thermo_c_long$Str_tr <- factor(thermo_c_long$Str_tr, levels = c("N2_ev", "N2_daf", "SKN-1_ev", "SKN-1_daf"))

#change names of Str_tr
levels(thermo_c_long$Str_tr) <- list('N2 ev' = 'N2_ev', 'N2 daf-2' = 'N2_daf', 'SKN-1 ev' = 'SKN-1_ev', 'SKN-1 daf-2' = 'SKN-1_daf')

#Create colour palette
palette <- c('#E64B35FF','#4DBBD5FF', '#00A087FF', '#3C5488FF')

#Age-specific reproduction plot
thermo_C_p <-ggplot(data=thermo_c_long, aes(x=factor(Day), y=value, group=Str_tr, color=Str_tr))+
  geom_jitter(alpha = 0.2, position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.5))+
  stat_summary(fun.data="mean_cl_boot", geom="errorbar", size = 1.2, width=0.0, position = position_dodge(0.5)) +
  stat_summary(fun.data="mean_cl_boot", geom="point", size = 3, position = position_dodge(0.5)) +
  stat_summary(fun.data="mean_cl_boot", geom="line",  size=1.2, position = position_dodge(0.5)) +
  theme_classic()+
  labs(y="Offspring number", x="Day")+
  labs(col="")+
  labs(title = 'Start 15\u00B0C')+   
  theme(axis.title.y = element_text(size=12))+
  theme(axis.title.x = element_text(size=12))+
  theme(axis.text= element_text(size = 12))+
  theme(legend.text = element_text(size = 12))+
  theme(plot.title = element_text(size =12, face = 'bold'))+
  scale_color_manual(values=palette, labels = c('N2 ev', N2_daf_label, skn_ev_label, skn_daf_label))+
  theme(legend.key.width = unit(0.5,"cm"))+
  coord_cartesian(ylim = c(0,175))+
  theme(legend.position = c(0.8, 0.9))

thermo_C_p


# Lifetime reproduction success

#Create dataframe containing Sum of total reproduction for each individual
totalrep_CF<-na.omit(as.data.frame.table(tapply(thermo_c_long$value,list(thermo_c_long$Strain, thermo_c_long$Treatment, thermo_c_long$Str_tr, thermo_c_long$ID),sum)))

#rename columns
names(totalrep_CF)<-c('Strain', 'Treatment', "Str_tr", "Replicate", "Totrep")

#Set factor levels for 
totalrep_CF$Str_tr <- factor(totalrep_CF$Str_tr, levels = c("N2 ev", "N2 daf-2", "SKN-1 ev", "SKN-1 daf-2"))

#linear model
LRS_CF2 <- lm(Totrep ~ Strain + Treatment, data = totalrep_CF)


LRS_CF_forest <- forest_emmeans(LRS_CF2, xlab = lrs_lab)
LRS_CF_forest

#change colours and labels
LRS_CF_forest <- LRS_CF_forest + 
  scale_colour_manual(values = c('#4DBBD5FF', '#00A087FF', '#3C5488FF'))+ 
  scale_y_discrete(labels =c(N2_daf_label, skn_ev_label, skn_daf_label))+ 
  theme(legend.position = 'none',
        axis.text = element_text(size = 12))

LRS_CF_forest


#calculate lambda
#Dataframe into long form to start fresh
thermo_c_long <- thermo_CF_rep %>% 
  pivot_longer(
    cols = c(`D1`,`D2`, `D3`, `D4`, `D5`,`D6`, `D7`,`D8`,`D9`), 
    names_to = "Day",
    values_to = "value")

#create a level order for 10 days
level_order <- c('D1', 'D2', 'D3', 'D4', 'D5','D6','D7','D8','D9')

#reorder levels so 10 is last
thermo_c_long$Day <- factor(thermo_c_long$Day, levels = level_order)


#spread data back out
thermo_c_wide <- spread(thermo_c_long, Day, value)
str(thermo_c_wide)

thermo_c_wide <- thermo_c_wide %>%
  dplyr::select(-c(Matricide))

thermo_c_wide <- na.omit(thermo_c_wide)

#calculate Lambda for each individual
L <- matrix(nrow = nrow(thermo_c_wide), ncol = 4)

for (i in 1:nrow(thermo_c_wide)){
  Les <- matrix(0, ncol = 11, nrow = 11)
  diag(Les[-1,]) <- rep(1, 10) # add the 1s for survival probability diagonally
  Fert <- c(0,0, as.numeric(as.vector(thermo_c_wide[i,][4:12]))) #the columns in data that has the reproductive counts
  Fert[is.na(Fert)] <- 0 #makes all NAs into 0s
  Les[1,] <- c(Fert)
  class(Les) <- "leslie.matrix"
  Lambda <- popbio::eigen.analysis(Les)$lambda1
  L[i, 1:4] <- c(paste0(thermo_c_wide$ID[i]), paste0(thermo_c_wide$Strain[i]), paste0(thermo_c_wide$Treatment[i]), Lambda)
  
}

#rename columns
colnames(L)<-c("ID", "Strain", "Treatment", "Lambda")

#make L a data frame
thermo_c_lam<-as.data.frame(L)

#make sure it's numeric
thermo_c_lam$Lambda<-as.numeric(as.character(thermo_c_lam$Lambda))
thermo_c_lam$Treatment <- as.factor(thermo_c_lam$Treatment)

#linear model
thermo_c_lam_m2 <- lm(Lambda ~ Strain + Treatment, data = thermo_c_lam)

Lam_CF_forest <- forest_emmeans(thermo_c_lam_m2, xlab = lam_lab)
Lam_CF_forest

#change colours and labels
Lam_CF_forest <- Lam_CF_forest + 
  scale_colour_manual(values = c('#4DBBD5FF', '#00A087FF', '#3C5488FF'))+
  scale_y_discrete(labels =c(N2_daf_label, skn_ev_label, skn_daf_label))

Lam_CF_forest

CF_Lam_LRS <- ggarrange(LRS_CF_forest+theme(plot.margin = unit(c(0, 0.5, 0.25, 0.45), 'cm')),
                        Lam_CF_forest+theme(plot.margin = unit(c(0, 0.5, 0.25, 0.42), 'cm')),
                        nrow = 2, labels = c('C','E'))
CF_Lam_LRS

CF_rep_all <- ggarrange(thermo_C_p+theme(plot.margin = unit(c(0, 0.5,0, 0.45), 'cm')), CF_Lam_LRS, nrow = 2, labels = c('A',''), heights = c(1, 1), common.legend = T)
CF_rep_all

```

## Lifespan plots
```{r}

#Create a combined variable for strain and treatment since forest plot can't do interactions
thermo_CF_LS <- thermo_CF_LS %>%
  unite(Str_tr, c('Strain', 'Treatment.ID'), remove = F)



#Set levels
thermo_CF_LS$Tr_str <- factor(thermo_CF_LS$Str_tr, levels = c("N2_ev", "N2_daf", "SKN-1_ev", "SKN-1_daf"))

#Set as factor
thermo_CF_LS$Str_tr <- as.factor(thermo_CF_LS$Str_tr)

#Set N2_ev as baseline level
thermo_CF_LS$Str_tr <- relevel(thermo_CF_LS$Str_tr, ref = 'N2_ev')

#Exclude unnatural causes of mortality, L = loss, W = wall, E = explosion
thermo_CF_LS <- thermo_CF_LS %>%
  filter(thermo_CF_LS$Cause != 'L' & thermo_CF_LS$Cause !='W' & thermo_CF_LS$Cause != 'E')


#Create survival object for plotting Kaplan Meier curves
surv_CF<-survfit(Surv(Age,Event)~Str_tr,data=thermo_CF_LS)

#Survival curve plot
THC_LS_plot <-ggsurvplot(surv_CF, ylab="Survival probability", data = thermo_CF_LS, size= 0.8, font.ylab= 14, font.xlab= 14, legend = c(0.3, 0.4), legend.title = "", title = "Start 15\u00B0C", censor = FALSE, xlab = "Day", xlim=c(0,40), break.time.by = 5, palette = palette, position= position_dodge(0.9), legend.labs=c("N2 ev", "N2 daf-2", "SKN-1 ev", "SKN-1 daf-2"),font.tickslab = c(12),  font.legend = c(13), font.title = c(14))
THC_LS_plot

THC_LS_plot <- THC_LS_plot$plot + scale_color_manual(labels = c(  'N2 ev',N2_daf_label,skn_ev_label,skn_daf_label), values = c('#E64B35FF', '#4DBBD5FF','#00A087FF','#3C5488FF'))+theme(plot.title = element_text(face = 'bold'))

THC_LS_plot
#Create dataframe excluding matricides
thermo_CF_LSNM <- thermo_CF_LS %>%
  filter(thermo_CF_LS$Cause != 'M')


#Create survival object
survCF_NM<-survfit(Surv(Age,Event)~Str_tr,data=thermo_CF_LSNM)

#Create survival curves
LS_plot_matcen_THC <-ggsurvplot(survCF_NM, ylab="Survival probability", data = thermo_CF_LSNM, size= 0.8, font.ylab= 12, font.xlab= 12, legend = c(0.3, 0.4), legend.title = "", palette = palette, title = "Start 15\u00B0C", censor = FALSE, xlab = "Day", xlim=c(0,40), break.time.by = 5, position= position_dodge(0.9), legend.labs=c("N2 ev", "N2 daf-2", "SKN-1 ev", "SKN-1 daf-2"), font.tickslab = c(12), font.legend = c(12), font.title = c(12))
LS_plot_matcen_THC


LS_plot_matcen_THC <- LS_plot_matcen_THC$plot + scale_color_manual(labels = c(  'N2 ev',N2_daf_label,skn_ev_label,skn_daf_label), values = c('#E64B35FF', '#4DBBD5FF','#00A087FF','#3C5488FF'))+theme(plot.title = element_text(face = 'bold'))

LS_plot_matcen_THC
#Sometimes NAs are produced creating an additional non-existent factor level - omit NAs first
thermo_CF_LS <- na.omit(thermo_CF_LS)
thermo_CF_LSNM <- na.omit(thermo_CF_LSNM)

#Drop non-existent factor level - or cox models get confused
thermo_CF_LS$Str_tr <- droplevels(thermo_CF_LS$Str_tr)
levels(thermo_CF_LS$Str_tr)
thermo_CF_LSNM$Str_tr <- droplevels(thermo_CF_LSNM$Str_tr)

#Mixed effects cox proportional hazards model to test effect of treatment on survival probability
CF_cox <- coxme(Surv(Age, Event) ~ Strain * Treatment.ID + (1|Plate.ID), data = thermo_CF_LS)
summary(CF_cox)
Anova(CF_cox, type = 'III')

#Test model 
cox.zph(CF_cox)#Strain * Treatment is fine
plot(cox.zph(CF_cox))


CF_cox_forest <- forest_coxme(CF_cox, xlab = "Log hazard ratio")
CF_cox_forest

CF_cox_forest <- CF_cox_forest + scale_color_manual(values = c('#4DBBD5FF','#00A087FF','#3C5488FF'))+ scale_y_discrete(labels = c(N2_daf_label,skn_ev_label,skn_daf_label))


CF_cox_forest


#Mixed effects cox proportional hazards model to test effect of treatment on survival prob
CF_NM_cox <- coxme(Surv(Age, Event) ~ Strain * Treatment.ID + (1|Plate.ID), data = thermo_CF_LSNM)
summary(CF_NM_cox)
Anova(CF_NM_cox, type = 'III')

#test model
plot(cox.zph(CF_NM_cox))
cox.zph(CF_NM_cox)#close p-values - combined with plot acceptable 

CF_NM_cox_forest <- forest_coxme(CF_NM_cox, xlab = "Log hazard ratio")
CF_NM_cox_forest

CF_NM_cox_forest <- CF_NM_cox_forest + scale_color_manual(values = c('#4DBBD5FF','#00A087FF','#3C5488FF')) + scale_y_discrete(labels = c(N2_daf_label,skn_ev_label,skn_daf_label))


CF_NM_cox_forest

```


# Hot first thermocycling

Reproduction and lifespan assays following 24 hour thermo-cycling between 15 and 25C with 25C as starting temperature across four treatments: SKN-1 mutant + daf-2 RNAi, SKN-1 mutant + ev (control), N2 + daf-2 RNAi, N2 + ev

## Reproduction
```{r}

#Put dataframe into long form
thermo_h_long <- thermo_HF_rep %>% 
  pivot_longer(
    cols = c(`D1`,`D2`, `D3`, `D4`, `D5`,`D6`, `D7`,`D8`,`D9`), 
    names_to = "Day",
    values_to = "value")

#Keep only columns that are necessary
thermo_h_long <- thermo_h_long %>%
  dplyr::select(c('ID','Strain', 'Treatment','Day','value'))

#Get rid of NAs
thermo_h_long <- na.omit(thermo_h_long)

#Create a combined strain and treatment column for plotting
thermo_h_long <- thermo_h_long %>%
  unite(Str_tr, c('Strain', 'Treatment'), remove = F)

#Set factor levels for Strain x Treatment
thermo_h_long$Str_tr <- factor(thermo_h_long$Str_tr, levels = c("N2_ev", "N2_daf", "SKN-1_ev", "SKN-1_daf"))

#change names of Str_tr
levels(thermo_h_long$Str_tr) <- list('N2 ev' = 'N2_ev', 'N2 daf-2' = 'N2_daf', 'SKN-1 ev' = 'SKN-1_ev', 'SKN-1 daf-2' = 'SKN-1_daf')

#Age-specific reproduction plot
thermo_H_p <-ggplot(data=thermo_h_long, aes(x=factor(Day), y=value, group=Str_tr, color=Str_tr))+
  geom_jitter(alpha = 0.2, position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.5))+
  stat_summary(fun.data="mean_cl_boot", geom="errorbar", size = 1, width=0.0, position = position_dodge(0.5)) +
  stat_summary(fun.data="mean_cl_boot", geom="point", size = 2, position = position_dodge(0.5)) +
  stat_summary(fun.data="mean_cl_boot", geom="line",  size=1, position = position_dodge(0.5)) +
  theme_classic()+
  labs(y="Offspring number", x="Day")+
  labs(col="")+
  labs(title = 'Start 25\u00B0C')+
  theme(plot.title = element_text(size = 12, face = 'bold'))+
  theme(axis.title.y = element_text(size=12))+
  theme(axis.title.x = element_text(size=12))+
  theme(axis.text= element_text(size = 12))+
  theme(legend.text = element_text(size = 12))+
  scale_color_manual(values=palette, labels = c('N2 ev', N2_daf_label, skn_ev_label, skn_daf_label))+
  theme(legend.key.width = unit(0.5,"cm"))+
  coord_cartesian(ylim = c(0,225))+
  theme(legend.position = c(0.8, 0.9))

thermo_H_p

#Calculate total reproduction over experiment per individual
totalrep_HF<-na.omit(as.data.frame.table(tapply(thermo_h_long$value,list(thermo_h_long$Strain, thermo_h_long$Treatment, thermo_h_long$Str_tr, thermo_h_long$ID),sum)))

#rename columns
names(totalrep_HF)<-c('Strain','Treatment',"Str_tr", "Replicate", "Totrep")

#change factor levels
totalrep_HF$Str_tr <- factor(totalrep_HF$Str_tr, levels = c("N2 ev", "N2 daf-2", "SKN-1 ev", "SKN-1 daf-2"))



totalrep_HF$Treatment <- relevel(totalrep_HF$Treatment, ref = 'ev')

#linear model
LRS_HF_m2 <- lm(Totrep ~ Treatment + Strain, data = totalrep_HF)


LRS_HF_forest <- forest_emmeans(LRS_HF_m2, xlab = lrs_lab)


LRS_HF_forest <- LRS_HF_forest + 
  scale_colour_manual(values = c('#4DBBD5FF', '#00A087FF', '#3C5488FF'))+ 
  scale_y_discrete(labels =c(N2_daf_label, skn_ev_label, skn_daf_label))

LRS_HF_forest

#Dataframe into long form to start fresh
thermo_h_long <- thermo_HF_rep %>% 
  pivot_longer(
    cols = c(`D1`,`D2`, `D3`, `D4`, `D5`,`D6`, `D7`,`D8`,`D9`), 
    names_to = "Day",
    values_to = "value")

#create a level order for 10 days
level_order <- c('D1', 'D2', 'D3', 'D4', 'D5','D6','D7','D8','D9')

#reorder levels so 10 is last
thermo_h_long$Day <- factor(thermo_h_long$Day, levels = level_order)


#spread data back out
thermo_h_wide <- spread(thermo_h_long, Day, value)
str(thermo_h_wide)

thermo_h_wide <- thermo_h_wide %>%
  dplyr::select(-c(Matricide))

thermo_h_wide <- na.omit(thermo_h_wide)

#calculate Lambda for each individual
L <- matrix(nrow = nrow(thermo_h_wide), ncol = 4)

for (i in 1:nrow(thermo_h_wide)){
  Les <- matrix(0, ncol = 11, nrow = 11)
  diag(Les[-1,]) <- rep(1, 10) # add the 1s for survival probability diagonally
  Fert <- c(0,0, as.numeric(as.vector(thermo_h_wide[i,][4:12]))) #the columns in data that has the reproductive counts
  Fert[is.na(Fert)] <- 0 #makes all NAs into 0s
  Les[1,] <- c(Fert)
  class(Les) <- "leslie.matrix"
  Lambda <- popbio::eigen.analysis(Les)$lambda1
  L[i, 1:4] <- c(paste0(thermo_h_wide$ID[i]), paste0(thermo_h_wide$Strain[i]), paste0(thermo_h_wide$Treatment[i]), Lambda)
  
}

#rename columns
colnames(L)<-c("ID", "Strain", "Treatment", "Lambda")

#make L a data frame
thermo_h_lam<-as.data.frame(L)

#make sure it's numeric
thermo_h_lam$Lambda<-as.numeric(as.character(thermo_h_lam$Lambda))
thermo_h_lam$Treatment <- as.factor(thermo_h_lam$Treatment)


#linear model
thermo_h_lam_m1 <- lm(Lambda ~ Strain * Treatment, data = thermo_h_lam)

Lam_HF_forest <- forest_emmeans(thermo_h_lam_m1, xlab = lam_lab)

#change colours and labels
Lam_HF_forest <- Lam_HF_forest + 
  scale_colour_manual(values = c('#4DBBD5FF', '#00A087FF', '#3C5488FF'))+ 
  scale_y_discrete(labels =c(N2_daf_label, skn_ev_label, skn_daf_label))


Lam_HF_forest




HF_Lam_LRS <- ggarrange(LRS_HF_forest+theme(plot.margin = unit(c(0, 0.5, 0.25, 0.45), 'cm')),
                        Lam_HF_forest+theme(plot.margin = unit(c(0, 0.5, 0.25, 0.45), 'cm')),
                        nrow = 2, labels = c('D','F'))
HF_Lam_LRS

HF_rep_all <- ggarrange(thermo_H_p+theme(plot.margin = unit(c(0, 0.5,0, 0.45), 'cm')), HF_Lam_LRS, nrow = 2, labels = c('B',''), heights = c(1, 1), common.legend = T)
HF_rep_all

```

## Lifespan
```{r}
#Combine Strain and Treatment into one variable for plotting
thermo_HF_LS <- thermo_HF_LS %>%
  unite(Str_tr, c('Strain', 'Treatment.ID'), remove = F)

#Make sure it's a factor
thermo_HF_LS$Str_tr <- as.factor(thermo_HF_LS$Str_tr)

#Set N2 Ev as reference level
thermo_HF_LS$Str_tr <- relevel(thermo_HF_LS$Str_tr, ref = 'N2_ev')

#FIlter out unnatural deaths (Loss, walling, explosions)
thermo_HF_LS <- thermo_HF_LS %>%
  filter(thermo_HF_LS$Cause != 'L' & thermo_HF_LS$Cause !='W' & thermo_HF_LS$Cause != 'E')

#name factor levels of Str_tr
thermo_HF_LS$Str_tr <- factor(thermo_HF_LS$Str_tr, levels = c("N2_ev", "N2_daf", "SKN-1_ev", "SKN-1_daf"))

#Check what Age is classed as
class(thermo_HF_LS$Age)

#Change to numeric
thermo_HF_LS$Age <- as.numeric(thermo_HF_LS$Age)

#Create survival object
survHF<-survfit(Surv(Age,Event)~Str_tr,data=thermo_HF_LS)

#Create a binary variable for matricides (1 = matricide, 0 = no matricide) to analyse with logistic regression
thermo_HF_LS$Matricide <- as.numeric(thermo_HF_LS$Cause == 'M')


#binomial glmer for matricide occurrence (logistic regression)
mat_mod <- glmer(Matricide ~ Str_tr + (1|Plate.ID), data = thermo_HF_LS, family = 'binomial')
summary(mat_mod)

#Create colour palette
palette <- c('#E64B35FF','#4DBBD5FF', '#00A087FF', '#3C5488FF')

#Survival curve
THH_LS_plot <-ggsurvplot(survHF, ylab="Survival probability", data = thermo_HF_LS, size= 0.8, font.ylab= 14, font.xlab= 14, legend = c(0.8, 0.9), legend.title = "", title = "Start 25\u00B0C", censor = FALSE, xlab = "Day", xlim=c(0,40), break.time.by = 5, palette = palette, position= position_dodge(0.9), legend.labs=c("N2 ev", "N2 daf-2", "SKN-1 ev", "SKN-1 daf"),font.tickslab = c(12),  font.legend = c(13), font.title = c(14))
THH_LS_plot

THH_LS_plot <- THH_LS_plot$plot + scale_color_manual(labels = c(  'N2 ev',N2_daf_label,skn_ev_label,skn_daf_label), values = c('#E64B35FF', '#4DBBD5FF','#00A087FF','#3C5488FF')) +theme(plot.title = element_text(face = 'bold'))

THH_LS_plot

??ggsurvplot

#Exclude matricides
thermo_HF_LSNM <- thermo_HF_LS %>%
  filter(thermo_HF_LS$Cause != 'M')


#Create survival object for matricide censored data
survHF_NM<-survfit(Surv(Age,Event)~Str_tr,data=thermo_HF_LSNM)

#Create survival curves
LS_plot_matcen_THH <-ggsurvplot(survHF_NM, ylab="Survival probability", data = thermo_HF_LSNM, size= 0.8, font.ylab= 12, font.xlab= 12, legend = c(0.2, 0.2), legend.title = "", palette = palette, title = "Start 25\u00B0C", censor = FALSE, xlab = "Day", xlim=c(0,40), break.time.by = 5, position= position_dodge(0.9), legend.labs=c("N2 ev", "N2 daf-2", "SKN-1 ev", "SKN-1 daf-1"), font.tickslab = c(12), font.legend = c(12), font.title = c(12))
LS_plot_matcen_THH

LS_plot_matcen_THH <- LS_plot_matcen_THH$plot + scale_color_manual(labels = c(  'N2 ev',N2_daf_label,skn_ev_label,skn_daf_label), values = c('#E64B35FF', '#4DBBD5FF','#00A087FF','#3C5488FF'))+theme(plot.title = element_text(face = 'bold'))

LS_plot_matcen_THH

#Sometimes NAs are produced creating an additional non-existent factor level - omit NAs first
thermo_HF_LS <- na.omit(thermo_HF_LS)
thermo_HF_LSNM <- na.omit(thermo_HF_LSNM)

#Drop non-existent factor level - or cox models get confused
thermo_HF_LS$Str_tr <- droplevels(thermo_HF_LS$Str_tr)
levels(thermo_HF_LS$Str_tr)
thermo_HF_LSNM$Str_tr <- droplevels(thermo_HF_LSNM$Str_tr)

#Mixed effects cox proportional hazards model to test effect of treatment on survival probability
HF_cox <- coxme(Surv(Age, Event) ~ Strain * Treatment.ID + (1|Plate.ID), data = thermo_HF_LS)
summary(HF_cox)
Anova(HF_cox, type = 'III')

#Test model 
cox.zph(HF_cox)#check with EH analysis

#forest plot
HF_cox_forest <- forest_coxme(HF_cox, xlab = "Log hazard ratio")
HF_cox_forest

#add colours and labels
HF_cox_forest <- HF_cox_forest + scale_color_manual(values = c('#4DBBD5FF','#00A087FF','#3C5488FF')) + scale_y_discrete(labels = c(N2_daf_label,skn_ev_label,skn_daf_label))

HF_cox_forest

#Mixed effects cox proportional hazards model to test effect of treatment on survival prob
HF_NM_cox <- coxme(Surv(Age, Event) ~ Strain * Treatment.ID + (1|Plate.ID), data = thermo_HF_LSNM)
summary(HF_NM_cox)
Anova(HF_NM_cox, type = 'III')
#test model
plot(cox.zph(HF_NM_cox))
cox.zph(HF_NM_cox)#compare to an EH analysis


#forest plot
HF_NM_cox_forest <- forest_coxme(HF_NM_cox, xlab = "Log hazard ratio")
HF_NM_cox_forest

HF_NM_cox_forest <- HF_NM_cox_forest + scale_color_manual(values = c('#4DBBD5FF','#00A087FF','#3C5488FF')) + scale_y_discrete(labels = c(N2_daf_label,skn_ev_label,skn_daf_label))

HF_NM_cox_forest

```

Combine thermocycle reproduction lifespan plots
##need to run previous code chunks to combine
```{r}

#combine reproduction plots
thermo_rep_all <- ggarrange(CF_rep_all, HF_rep_all, ncol = 2, common.legend = T)
thermo_rep_all

#(plot = thermo_rep_all, 'thermo_rep_all.png', width = 9, height = 7, bg = 'white')

#combine lifespan plots
thermo_NM_LS_all <- ggarrange(LS_plot_matcen_THC,CF_NM_cox_forest, LS_plot_matcen_THH,  HF_NM_cox_forest, ncol = 1, nrow = 4, common.legend = T, heights = c(1.75, 1), labels = 'AUTO')
thermo_NM_LS_all

thermo_NM_LS_all_2 <- ggarrange(LS_plot_matcen_THC,LS_plot_matcen_THH,CF_NM_cox_forest,   HF_NM_cox_forest, ncol = 2, nrow = 2, common.legend = T, heights = c(1.75, 1), labels = 'AUTO')
thermo_NM_LS_all_2


#ggsave(plot = thermo_NM_LS_all_2, 'thermo_NM_LS_all_2.png', height = 7, width = 10, bg = 'white')
#ggsave(plot = thermo_NM_LS_all, 'thermo_NM_LS_all.png', height = 12, width = 6, bg = 'white')
```


# SKN-1 in 15C
Lifespan assays in SKN-1 mutants and N2s with daf-2 or ev (control) RNAi treatments in 15C, including a treatment with a single 24-hour cycle of 25C at ~ L2 stage
##Lifespan
```{r}

#Include only 15C data
SKN_15 <- SKN_15_25_LS %>%
  filter(Temperature != '25' & Temperature != '25F')

#Unite strain and treatment for plotting
SKN_15 <- SKN_15 %>%
  unite(Str_tr, 'Strain','Treatment.ID', remove = F)

#Unite strain, treatment and temperature for plotting
SKN_15 <- SKN_15 %>%
  unite(Str_tr_temp, 'Str_tr', 'Temperature', remove = F)

#Make sure it's a factor
SKN_15$Str_tr_temp <- as.factor(SKN_15$Str_tr_temp)

#Set reference level to N2 Ev
SKN_15$Str_tr_temp <- relevel(SKN_15$Str_tr_temp, ref = 'N2_ev_15')

#Name levels
SKN_15$Str_tr_temp <- factor(SKN_15$Str_tr_temp, levels = c('N2_ev_15','N2_daf_15','SKN_ev_15','SKN_daf_15','SKN_daf_15F'))

#Create survival object
surv_SKN15<-survfit(Surv(Age,Event)~Str_tr + Temperature,data=SKN_15)

#Create survival curves
SKN_15_plot <-ggsurvplot(surv_SKN15, ylab="Survival probability", data = SKN_15, size= 0.8, font.ylab= 12, font.xlab= 12, legend = c(0.2, 0.2), legend.title = "", title = "15°C",palette = c('#4DBBD5FF','#E64B35FF' ,'#3C5488FF','#F39B7FFF','#00A087FF'), censor = FALSE, xlab = "Day", xlim=c(0,40), break.time.by = 5, position= position_dodge(0.9), font.tickslab = c(12), font.legend = c(12), font.title = c(12))

SKN_15_plot

SKN_15_plot <- SKN_15_plot$plot + scale_color_manual(labels = c(N2_daf_label,  'N2 ev',skn_daf_label,skn_cycle_label,skn_ev_label), values = c('#4DBBD5FF','#E64B35FF' ,'#3C5488FF','#F39B7FFF','#00A087FF'))+theme(plot.title = element_text(face = 'bold'))

SKN_15_plot

#Exclude matricides
SKN_15_NM <- SKN_15 %>%
  filter(Cause != 'M')


#Create survival objet
surv_SKN15_NM<-survfit(Surv(Age,Event)~Str_tr + Temperature,data=SKN_15_NM)

#Create survival curves
SKN_15_NM_plot <-ggsurvplot(surv_SKN15_NM, ylab="Survival probability", data = SKN_15_NM, size= 0.8, font.ylab= 12, font.xlab= 12, legend = c(0.2, 0.2), legend.title = "", title = "15°C", censor = FALSE, xlab = "Day", xlim=c(0,40), break.time.by = 5, position= position_dodge(0.9), font.tickslab = c(12), palette = c('#4DBBD5FF','#E64B35FF', '#3C5488FF','#F39B7FFF','#00A087FF'), font.legend = c(12),font.title = c(12))

SKN_15_NM_plot

SKN_15_NM_plot <- SKN_15_NM_plot$plot + scale_color_manual(labels = c(N2_daf_label,  'N2 ev',skn_daf_label,skn_cycle_label,skn_ev_label), values = c('#4DBBD5FF','#E64B35FF' ,'#3C5488FF','#F39B7FFF','#00A087FF'))+theme(plot.title = element_text(face = 'bold'))

SKN_15_NM_plot
#Analyse two strains individually due to additionally treatment of thermocycling during development in SKN-1 only
#FIlter only SKNs 
SKN_only_15 <- SKN_15 %>%
  filter(Strain == 'SKN')

str(SKN_only_15)

#Check levels of Strain
SKN_only_15$Strain <- as.factor(SKN_only_15$Strain)
levels(SKN_only_15$Strain)

#Check levels of temperature
levels(SKN_only_15$Temperature)

#Need to drop levels - still thinks it has 15C treatments 
SKN_only_15$Temperature <- droplevels(SKN_only_15$Temperature)
levels(SKN_only_15$Temperature)

#Check levels of Treatment
SKN_only_15$Treatment.ID <- as.factor(SKN_only_15$Treatment.ID)
levels(SKN_only_15$Treatment.ID)
SKN_only_15$Treatment.ID <- droplevels(SKN_only_15$Treatment.ID)

#Combine Treatment and temperature because of imbalance
SKN_only_15 <- SKN_only_15 %>%
  unite(Tr_temp, 'Treatment.ID','Temperature', remove = F)

#set ev as reference level
SKN_only_15$Tr_temp <- as.factor(SKN_only_15$Tr_temp)
SKN_only_15$Tr_temp <- relevel(SKN_only_15$Tr_temp, ref = 'ev_15')

#Coxme model with combined effect of treatment and temperature to account for developmental thermocycle treatment
cox_SKN_15_only <- coxme(Surv(Age, Event) ~ Tr_temp + (1|Plate.ID), data = SKN_only_15)
summary(cox_SKN_15_only)
Anova(cox_SKN_15_only, type = 'II')

#test model
cox.zph(cox_SKN_15_only)# OK! p = 0.18

#Matricides censored
SKN_only_15_NM <- SKN_only_15 %>%
  filter(Cause !='M')

#Coxme model with combined effect of treatment and temperature to account for developmental thermocycle treatment with matricides censored
cox_SKN_15_only_NM <- coxme(Surv(Age, Event) ~ Tr_temp + (1|Plate.ID), data = SKN_only_15_NM)
summary(cox_SKN_15_only_NM)
Anova(cox_SKN_15_only_NM, type = 'II')
#test model
cox.zph(cox_SKN_15_only_NM)# global p = 0.41, OK!


#Filter out N2s
N2_only_15 <- SKN_15 %>%
  filter(Strain == 'N2')

#Set ev as reference level
N2_only_15$Treatment.ID <- as.factor(N2_only_15$Treatment.ID)
N2_only_15$Treatment.ID <- relevel(N2_only_15$Treatment.ID, ref = 'ev')

#Coxme model with treatment as factor
cox_N2_15_only <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = N2_only_15)
summary(cox_N2_15_only)
Anova(cox_N2_15_only, type = 'II')

#test model
cox.zph(cox_N2_15_only)#OK p = 0.79

#Censor matricides
N2_only_15_NM <- N2_only_15 %>%
  filter(Cause !='M')

#coxme model with treatment as factor with matricides censored
cox_N2_15_only_NM <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = N2_only_15_NM)
summary(cox_N2_15_only_NM)
Anova(cox_N2_15_only_NM, type = 'II')
#test model
cox.zph(cox_N2_15_only_NM)#p = 0.87 OK



#Forest plots with matricides
SKN_15_forest <- forest_cox_strain(
  cox_SKN_15_only,
  treatment = "Tr_temp",
  ref_treatment = "ev_15",
  xlab = "Log Hazard Ratio (vs ev)",
  treatment_labels = c("daf_15", "daf_15F"),
  vline_color = "#00A087FF"  # ordered display
)

SKN_15_forest

SKN_15_forest <- SKN_15_forest +
  scale_color_manual(values = c('#3C5488FF','#F39B7FFF'))+
  scale_y_discrete(labels = c(daf_label, daf_cyc_forest_lab))+
  theme(legend.position = 'none')+
  theme(plot.margin = margin(0, 0.5, 0, 0.5, unit = "cm"))+
  labs(title = skn_lab)

SKN_15_forest


N2_15_forest <- forest_cox_strain(
  cox_N2_15_only,
  treatment = "Treatment.ID",
  ref_treatment = "ev",
  xlab = "Log Hazard Ratio (vs ev)",
  treatment_labels = c("daf"),
  vline_color = "#E64B35FF"  # ordered display
)


N2_15_forest

N2_15_forest <- N2_15_forest +
  scale_color_manual(values = c('#4DBBD5FF'))+
  scale_y_discrete(labels = c(daf_label))+
  theme(legend.position = 'none')+
  theme(plot.margin = margin(0, 0.5, 0, 0.5, unit = "cm"))+
  labs(title = 'N2')

N2_15_forest


#Forest plots without matricides
SKN_15NM_forest <- forest_cox_strain(
  cox_SKN_15_only_NM,
  treatment = "Tr_temp",
  ref_treatment = "ev_15",
  xlab = "Log Hazard Ratio (vs ev)",
  treatment_labels = c("daf_15", "daf_15F"),
  vline_color = "#00A087FF"  # ordered display
)

SKN_15NM_forest

SKN_15NM_forest <- SKN_15NM_forest +
  scale_color_manual(values = c('#3C5488FF','#F39B7FFF'))+
  scale_y_discrete(labels = c(daf_label, daf_cyc_forest_lab))+
  theme(legend.position = 'none')+
  theme(plot.margin = margin(0, 0.5, 0, 0.5, unit = "cm"))+
  labs(title = skn_lab)

SKN_15NM_forest


N2_15NM_forest <- forest_cox_strain(
  cox_N2_15_only_NM,
  treatment = "Treatment.ID",
  ref_treatment = "ev",
  xlab = "Log Hazard Ratio (vs ev)",
  treatment_labels = c("daf"),
  vline_color = "#E64B35FF"  # ordered display
)


N2_15NM_forest

N2_15NM_forest <- N2_15NM_forest +
  scale_color_manual(values = c('#4DBBD5FF'))+
  scale_y_discrete(labels = c(daf_label))+
  theme(legend.position = 'none')+
  theme(plot.margin = margin(0, 0.5, 0, 0.5, unit = "cm"))+
  labs(title = 'N2')

N2_15NM_forest


```


# SKN-1 in 25C
Lifespan assays in SKN-1 mutants and N2s with daf-2 or ev (control) RNAi treatments in 25C, including a treatment with a single 24-hour cycle of 15C at ~ L2 stage
## Lifespan
```{r}

#Temperature as a factor
SKN_15_25_LS$Temperature <- as.factor(SKN_15_25_LS$Temperature)

#Get only 25C data
SKN_25 <- SKN_15_25_LS %>%
  filter(Temperature != '15' & Temperature != '15F')

#Combine Strain and treatment for plotting
SKN_25 <- SKN_25 %>%
  unite(Str_tr, 'Strain','Treatment.ID', remove = F)

#Also combine temperature with strain and treatment
SKN_25 <- SKN_25 %>%
  unite(Str_tr_temp, 'Str_tr', 'Temperature', remove = F)

#Rename factor levels
SKN_25$Str_tr_temp <- factor(SKN_25$Str_tr_temp, levels = c('N2_ev_25','N2_daf_25','SKN_ev_25','SKN_daf_25','SKN_daf_25F'))

#Ensure it's a factor
SKN_25$Str_tr_temp <- as.factor(SKN_25$Str_tr_temp)

#Set ev N2 with 25C only as reference level
SKN_25$Str_tr_temp <- relevel(SKN_25$Str_tr_temp, ref = 'N2_ev_25')

#Create survival object for plot
surv_SKN25<-survfit(Surv(Age,Event)~Str_tr + Temperature,data=SKN_25)

#Create survival curve
SKN_25_plot <-ggsurvplot(surv_SKN25, ylab="Survival probability", data = SKN_25, size= 0.8, font.ylab= 12, font.xlab= 12, legend = c(0.8, 0.8), legend.title = "", title = "25°C", palette = c('#4DBBD5FF','#E64B35FF' ,'#3C5488FF','#F39B7FFF','#00A087FF'),censor = FALSE, xlab = "Day", xlim=c(0,30), break.time.by = 5, position= position_dodge(0.9), font.tickslab = c(12), font.legend = c(12),font.title = c(12))

SKN_25_plot

SKN_25_plot <- SKN_25_plot$plot + scale_color_manual(labels = c(N2_daf_label,  'N2: ev',skn_daf_label,skn_cycle_label,skn_ev_label), values = c('#4DBBD5FF','#E64B35FF' ,'#3C5488FF','#F39B7FFF','#00A087FF'))+theme(plot.title = element_text(face = 'bold'))

SKN_25_plot

#Exclude matricides
SKN_25_NM <- SKN_25 %>%
  filter(Cause != 'M')

#Create survival object
surv_SKN25_NM<-survfit(Surv(Age,Event)~Str_tr + Temperature,data=SKN_25_NM)

#Create survival curve
SKN_25_NM_plot <-ggsurvplot(surv_SKN25_NM, ylab="Survival probability", data = SKN_25_NM, size= 0.8, font.ylab= 12, font.xlab= 12, legend = c(0.8, 0.8), legend.title = "", palette = c('#4DBBD5FF','#E64B35FF' ,'#3C5488FF','#F39B7FFF','#00A087FF'), title = "25°C",  censor = FALSE, xlab = "Day", xlim=c(0,30), break.time.by = 5, position= position_dodge(0.9), font.tickslab = c(12), font.legend = c(12),font.title = c(12)) 
#legend.labs = c('N2 daf-2', 'N2 ev','SKN-1 daf-2','SKN-1 daf-2 + cycle','SKN-1 ev'))
SKN_25_NM_plot
SKN_25_NM_plot <- SKN_25_NM_plot$plot + scale_color_manual(labels = c(N2_daf_label,  'N2: ev',skn_daf_label,skn_cycle_label,skn_ev_label), values = c('#4DBBD5FF','#E64B35FF' ,'#3C5488FF','#F39B7FFF','#00A087FF'))+theme(plot.title = element_text(face = 'bold'))

SKN_25_NM_plot


#Make sure Str-tr_temp is a factor
SKN_25_NM$Str_tr_temp <- as.factor(SKN_25_NM$Str_tr_temp)

#Reset baseline to ev N2 in 25 only
SKN_25_NM$Str_tr_temp <- relevel(SKN_25_NM$Str_tr_temp, ref = 'N2_ev_25')


#Analyse within strains becaues of additional treatment only in SKN-1

#FIlter only SKNs 
SKN_only_25 <- SKN_25 %>%
  filter(Strain == 'SKN')

str(SKN_only_25)

#Check levels of Strain
SKN_only_25$Strain <- as.factor(SKN_only_25$Strain)
levels(SKN_only_25$Strain)

#Check levels of temperature
levels(SKN_only_25$Temperature)

#Need to drop levels - still thinks it has 15C treatments 
SKN_only_25$Temperature <- droplevels(SKN_only_25$Temperature)
levels(SKN_only_25$Temperature)

#Check levels of Treatment
SKN_only_25$Treatment.ID <- as.factor(SKN_only_25$Treatment.ID)
levels(SKN_only_25$Treatment.ID)
SKN_only_25$Treatment.ID <- droplevels(SKN_only_25$Treatment.ID)

#Combine Treatment and temperature because of imbalance
SKN_only_25 <- SKN_only_25 %>%
  unite(Tr_temp, 'Treatment.ID','Temperature', remove = F)

#set ev as reference level
SKN_only_25$Tr_temp <- as.factor(SKN_only_25$Tr_temp)
SKN_only_25$Tr_temp <- relevel(SKN_only_25$Tr_temp, ref = 'ev_25')

#Coxme model with combined effect of treatment and temperature to account for developmental thermocycle treatment
cox_SKN_25_only <- coxme(Surv(Age, Event) ~ Tr_temp + (1|Plate.ID), data = SKN_only_25)
summary(cox_SKN_25_only)
Anova(cox_SKN_25_only, type = 'II')
#test model
cox.zph(cox_SKN_25_only)# global p = 0.004 - check with EH analysis

#Matricides censored
SKN_only_25_NM <- SKN_only_25 %>%
  filter(Cause !='M')

#Coxme model with combined effect of treatment and temperature to account for developmental thermocycle treatment with matricides censored
cox_SKN_25_only_NM <- coxme(Surv(Age, Event) ~ Tr_temp + (1|Plate.ID), data = SKN_only_25_NM)
summary(cox_SKN_25_only_NM)
Anova(cox_SKN_25_only_NM, type = 'II')

#test model
cox.zph(cox_SKN_25_only_NM)# global p = 0.054 OK


#Filter out N2s
N2_only_25 <- SKN_25 %>%
  filter(Strain == 'N2')

#Set ev as reference level
N2_only_25$Treatment.ID <- as.factor(N2_only_25$Treatment.ID)
N2_only_25$Treatment.ID <- relevel(N2_only_25$Treatment.ID, ref = 'ev')

#Coxme model with treatment as factor
cox_N2_only <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = N2_only_25)
summary(cox_N2_only)
Anova(cox_N2_only, type = 'II')

#test model
cox.zph(cox_N2_only)#OK p>0.05

#Censor matricides
N2_only_25_NM <- N2_only_25 %>%
  filter(Cause !='M')

#coxme model with treatment as factor with matricides censored
cox_N2_only_NM <- coxme(Surv(Age, Event) ~ Treatment.ID + (1|Plate.ID), data = N2_only_25_NM)
summary(cox_N2_only_NM)
Anova(cox_N2_only_NM, type = 'II')


#Forest plots with matricides
SKN_25_forest <- forest_cox_strain(
  cox_SKN_25_only,
  treatment = "Tr_temp",
  ref_treatment = "ev_25",
  xlab = "Log Hazard Ratio",
  treatment_labels = c("daf_25", "daf_25F"),
  vline_color = "#00A087FF"  # ordered display
)

SKN_25_forest

SKN_25_forest <- SKN_25_forest +
  scale_color_manual(values = c('#3C5488FF','#F39B7FFF'))+
  scale_y_discrete(labels = c(daf_label, daf_cyc_forest_lab))+
  theme(legend.position = 'none')+
  theme(plot.margin = margin(0, 0.5, 0, 0.5, unit = "cm"))+
  labs(title = skn_lab)

SKN_25_forest


N2_25_forest <- forest_cox_strain(
  cox_N2_only,
  treatment = "Treatment.ID",
  ref_treatment = "ev",
  xlab = "Log Hazard Ratio",
  treatment_labels = c("daf"),
  vline_color = "#E64B35FF"  # ordered display
)


N2_25_forest

N2_25_forest <- N2_25_forest +
  scale_color_manual(values = c('#4DBBD5FF'))+
  scale_y_discrete(labels = c(daf_label))+
  theme(legend.position = 'none')+
  theme(plot.margin = margin(0, 0.5, 0, 0.5, unit = "cm"))+
  labs(title = 'N2')

N2_25_forest


#Forest plots without matricides
SKN_25NM_forest <- forest_cox_strain(
  cox_SKN_25_only_NM,
  treatment = "Tr_temp",
  ref_treatment = "ev_25",
  xlab = "Log Hazard Ratio (vs ev)",
  treatment_labels = c("daf_25", "daf_25F"),
  vline_color = "#00A087FF"  # ordered display
)

SKN_25NM_forest

SKN_25NM_forest <- SKN_25NM_forest +
  scale_color_manual(values = c('#3C5488FF','#F39B7FFF'))+
  scale_y_discrete(labels = c(daf_label, daf_cyc_forest_lab))+
  theme(legend.position = 'none')+
  theme(plot.margin = margin(0, 0.5, 0, 0.5, unit = "cm"))+
  labs(title = skn_lab)

SKN_25NM_forest


N2_25NM_forest <- forest_cox_strain(
  cox_N2_only_NM,
  treatment = "Treatment.ID",
  ref_treatment = "ev",
  xlab = "Log Hazard Ratio (vs ev)",
  treatment_labels = c("daf"),
  vline_color = "#E64B35FF"  # ordered display
)


N2_25NM_forest <- N2_25NM_forest +
  scale_color_manual(values = c('#4DBBD5FF'))+
  scale_y_discrete(labels = c(daf_label))+
  theme(legend.position = 'none')+
  theme(plot.margin = margin(0, 0.5, 0, 0.5, unit = "cm"))+
  labs(title = 'N2')

N2_25NM_forest

#combine 15/25C forest plots with matricides
SKN_N2_15_forest <- ggarrange(N2_15_forest,
                              SKN_15_forest,
                              ncol = 2, labels = c('B','C'))
SKN_N2_15_forest


SKN_N2_25_forest <- ggarrange(N2_25_forest,
                              SKN_25_forest, 
                              ncol = 2, labels = c('E','F'))
SKN_N2_25_forest

#combine 15/25C forest plots without matricides
SKN_N2_15NM_forest <- ggarrange(N2_15NM_forest,
                              SKN_15NM_forest,
                              ncol = 2, labels = c('B','C'))
SKN_N2_15NM_forest <- SKN_N2_15NM_forest + theme(plot.margin = unit(c(0,0,1,0), 'cm'))
SKN_N2_15NM_forest

SKN_N2_25NM_forest <- ggarrange(N2_25NM_forest,
                              SKN_25NM_forest, 
                              ncol = 2, labels = c('E','F'))
SKN_N2_25NM_forest <- SKN_N2_25NM_forest + theme(plot.margin = unit(c(0,0,1,0), 'cm'))
SKN_N2_25NM_forest


```


Combine constant temperature lifespan plots
##need to run previous 2 code chunks to combine all plots
```{r}
SKN_15_25_all <- ggarrange(SKN_15_plot + 
                             theme(plot.title = element_text(face = 'bold', size = 15)),
                           SKN_N2_15_forest,
                           SKN_25_plot +
                             theme(plot.title = element_text(face = 'bold', size = 15)),
                           SKN_N2_25_forest, 
                           nrow = 4, heights = c(1.75, 1, 1.75, 1), 
                           common.legend = T, labels = c('A','','D',''))

SKN_15_25_all

ggsave("SKN_15_25_all.png", plot = SKN_15_25_all, bg = 'white',width = 7, height = 11, dpi = 300)


SKN_15_25_NM_all <- ggarrange(SKN_15_NM_plot + 
                             theme(plot.title = element_text(face = 'bold', size = 15)),
                           SKN_N2_15NM_forest,
                           SKN_25_NM_plot +
                             theme(plot.title = element_text(face = 'bold', size = 15)),
                           SKN_N2_25NM_forest, 
                           nrow = 4, heights = c(1.65, 1, 1.65, 1), 
                           common.legend = T, labels = c('A','','D',''))

SKN_15_25_NM_all

ggsave("SKN_15_25_NM_all.png", plot = SKN_15_25_NM_all, bg = 'white',width = 7, height = 11, dpi = 300)


```
